---
title: "CAF-1_Paper_Figure1_combined_script_ET"
author: "Eddie Torres"
date: "2025-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Set working environment
workingdir="C:/Users/eddie/Lewis_Lab_Research/GitHub"

#Set working directory to the correct location for working machine
knitr::opts_knit$set(root.dir = "workingdir")
```

```{r load libraries, echo=FALSE}
library("DESeq2")
library("ggplot2")
library("gplots")
library("ggrepel")
library("dplyr")
library("pheatmap")
library("grid")
library("corrplot")
library("RColorBrewer")
library("scales")
library("readr")
library("tidyr")
library("tidyverse")
library("scater")
library("reshape2")
library("grDevices")
library("tibble")
library("VennDiagram")
library("eulerr")
library("gridExtra")
library("scales")
```

#Code for Violin plot & heatmap
##DESeq not yet need because I am just averaging raw TPMs for these figures. Will need to run DESeq for volcano plots & Venn Diagrams

```{r read in counts file from featureCounts, echo=FALSE}
#Bring in table with un-normalized transcription counts; check.names is important if you have dashes in the gene names 
##row.names=1 command sets geneIDs as the row name

countdataInteractors <- read.table("../text_files/readcounts_All_CAF1paper.merged.txt", header=TRUE,stringsAsFactors=FALSE, row.names=1, check.names=FALSE, sep="\t")

#create matrix containing count data only; i.e. eliminate contig, start, stop, length fields
##convert count file to matrix. Note:: I initially tried to use as.matrix, but this didn't work. <- data.matrix(countfile) worked fine.

Interactors_countsOnly <- data.matrix(countdataInteractors[ ,6:ncol(countdataInteractors)])

##Run calculateTPM function - must load "scater" library
Interactors_tpm <- calculateTPM(Interactors_countsOnly, lengths = countdataInteractors[,5])

##convert to matrix again
allDataTPM  <- data.matrix(Interactors_tpm)

#rename columns with human readable sample names
##Note: THIS NEXT SECTION REQUIRED MANUAL INPUT INTO A SPREADSHEET OR TEXT EDITOR, this code only reads in the file that was already renamed

names <- read.table(file="names_CAF.txt" , sep="", row.names = 1, header = TRUE)
colnames(allDataTPM) <- names[,1]
```

```{r select data from mutants you want to plot and calculate mean TPM, echo=FALSE}
#Make new matrix
Averaged_Orderd_KO_data <- cbind(rowMeans(allDataTPM[,22:25], na.rm = TRUE),
                                 rowMeans(allDataTPM[,26:28], na.rm = TRUE),
                                 rowMeans(allDataTPM[,29:31], na.rm = TRUE),
                                 rowMeans(allDataTPM[,4:6], na.rm = TRUE),
                                 rowMeans(allDataTPM[,7:9], na.rm = TRUE),
                                 rowMeans(allDataTPM[,10:12], na.rm = TRUE),
                                 rowMeans(allDataTPM[,16:18], na.rm = TRUE),
                                 rowMeans(allDataTPM[,13:15], na.rm = TRUE),
                                 rowMeans(allDataTPM[,19:21], na.rm = TRUE),
                                 rowMeans(allDataTPM[,32:34], na.rm = TRUE),
                                 rowMeans(allDataTPM[,35:37], na.rm = TRUE),
                                 rowMeans(allDataTPM[,38:40], na.rm = TRUE),
                                 rowMeans(allDataTPM[,41:42], na.rm = TRUE))
#Rename columns
averageRowIDs=c("WT","set-7","cac-1","cac-2","cac-3","naf-1","naf-2","asf-1","ATRX","cac-1_cac-2","cac-1_suz-12","suz-12","ash-1")
colnames(Averaged_Orderd_KO_data) <- averageRowIDs

#Make subsetted dfs i will want for figures

Data_Chaperones = Averaged_Orderd_KO_data [,1:9]
Data_ash1 = Averaged_Orderd_KO_data [,c(1,3:5,13)]
Data_Doubles = Averaged_Orderd_KO_data [,c(1,12,3,11)]

```

```{r subseting data to include genes of interest, echo=FALSE} 
#subset data to sample only PRC2-target genes
##reading in geneIDs of PRC2 target genes (promoters and genes fully covered by K27)
Prc2targets <- read.table("../bed_files/H3K27me3_methylated_genes_FINAL.bed", header=FALSE, stringsAsFactors=FALSE, check.names=FALSE, sep="\t") 

Prc2targetTPM <- subset(allDataTPM, rownames(allDataTPM)%in%Prc2targets[,10])

#This gene gets left out becausei had to rename it in my bed file due to blanks. I need to add it back manually because the subsetting doesn't recognize the new name
lac <- allDataTPM[rownames(allDataTPM) == "laccase precursor", ]
Prc2targetTPM <- rbind(Prc2targetTPM, lac)

#rename rows of added gene back to original name
rownames(Prc2targetTPM)[rownames(Prc2targetTPM) == "lac"] <- "laccase precursor"

#Subset Data (n = 575)
AVERAGE_Prc2targetTPM <- subset(Data_Chaperones, rownames(Data_Chaperones)%in%rownames(Prc2targetTPM))
AVERAGE_Prc2targetTPM_Double <- subset(Data_Doubles, rownames(Data_Doubles)%in%rownames(Prc2targetTPM))

#Subset data to filter out non-repressed PRC2 targets regions that got through (cutting out any genes over 10 tpm in WT). Choice of 10 TPM cutoff was arbitrary
##These are likely genes whose promoters are not marked by H3K27me3, genes on the edge of K27 regions, and/or bivalent genes
AVERAGE_Prc2targetTPM <- subset(AVERAGE_Prc2targetTPM, (AVERAGE_Prc2targetTPM[,1] < 10))
AVERAGE_Prc2targetTPM_Double <- subset(AVERAGE_Prc2targetTPM_Double, (AVERAGE_Prc2targetTPM_Double[,1] < 10))

#writing "repressed" genes to a new bed file to report
repress_bed = subset(Prc2targets, Prc2targets$V10 %in% rownames(AVERAGE_Prc2targetTPM))
write.table(repress_bed, file = "H3K27me3_Repressed_Genes.bed", row.names = FALSE, sep="\t", quote = FALSE)

#resubet PRC2 targets after filtering
Prc2targetTPM_Chap <- rownames(Prc2targetTPM) %in% rownames(AVERAGE_Prc2targetTPM)
Prc2targetTPM_Double <- subset(Prc2targetTPM, rownames(Prc2targetTPM)%in%rownames(AVERAGE_Prc2targetTPM_Double))

#Add sudocount and log transform (if necessary)
AVERAGE_Prc2targetTPM <- AVERAGE_Prc2targetTPM + 1
AVERAGE_Prc2targetTPM <- log2(AVERAGE_Prc2targetTPM)
#
AVERAGE_Prc2targetTPM_Double <- AVERAGE_Prc2targetTPM_Double + 1
AVERAGE_Prc2targetTPM_Double <- log2(AVERAGE_Prc2targetTPM_Double)


#melt data to get it into a format ggplot can use (library "reshape2")
meltedAveragePRC2targetData <- melt(AVERAGE_Prc2targetTPM, value.name = 'Count',
                             varnames=c('GeneID', 'Sample'))

meltedAverageDoublePRC2targetData <- melt(AVERAGE_Prc2targetTPM_Double, value.name = 'Count',
                             varnames=c('GeneID', 'Sample'))

```

```{r statistics, echo=FALSE}
#I want to run statistical tests to determine the signifcance of the difference in expression b/w mutant strains and WT

AVERAGE_Prc2targetTPM_ALL <- subset(Averaged_Orderd_KO_data, rownames(Averaged_Orderd_KO_data)%in%rownames(Prc2targetTPM))

#Subset data to filter out non-repressed PRC2 targets regions that got through (cutting out any genes over 10 tpm in WT). Choice of 10 TPM cutoff was arbitrary
##These are likely genes whose promoters are not marked by H3K27me3, genes on the edge of K27 regions, and/or bivalent genes
AVERAGE_Prc2targetTPM_ALL <- subset(AVERAGE_Prc2targetTPM_ALL, (AVERAGE_Prc2targetTPM_ALL[,1] < 10))

#resubet PRC2 targets after filtering
Prc2targetTPM_ALL <- subset(Prc2targetTPM, rownames(Prc2targetTPM)%in%rownames(AVERAGE_Prc2targetTPM_ALL))

#Add sudocount and log transform (if necessary)
AVERAGE_Prc2targetTPM_ALL <- AVERAGE_Prc2targetTPM_ALL + 1
AVERAGE_Prc2targetTPM_ALL <- log2(AVERAGE_Prc2targetTPM_ALL)

## First, I make histogram to visualize distribution of data. we'll use cac-1 as a comparison to WT
# hist <- as.data.frame(AVERAGE_Prc2targetTPM_ALL)
# 
# long_hist <- hist %>%
#   select(WT, `cac-1`) %>%
#   pivot_longer(cols = everything(), names_to = "Strain", values_to = "TPM")
# 
# # Plot overlapping histograms
# ggplot(long_hist, aes(x = TPM, fill = Strain)) +
#   geom_histogram(position = "identity", alpha = 0.5, bins = 15, color = "black") +
#   theme_minimal() +
#   labs(title = "TPM Distribution in WT and cac-1",
#        x = "TPM",
#        y = "Frequency") +
#   scale_fill_manual(values = c("WT" = "steelblue", "cac-1" = "tomato"))

#Histogram shows non-normal distribution (skewed right), so proceed with wilcox test instead of t test
##running tests
set7_stat <- wilcox.test(hist$`set-7`, hist$WT, alternative = "greater")
cac1_stat <- wilcox.test(hist$`cac-1`, hist$WT, alternative = "greater")
cac2_stat <- wilcox.test(hist$`cac-2`, hist$WT, alternative = "greater")
cac3_stat <- wilcox.test(hist$`cac-3`, hist$WT, alternative = "greater")
naf1_stat <- wilcox.test(hist$`naf-1`, hist$WT, alternative = "greater")
naf2_stat <- wilcox.test(hist$`naf-2`, hist$WT, alternative = "greater")
asf1_stat <- wilcox.test(hist$`asf-1`, hist$WT, alternative = "greater")
atrx_stat <- wilcox.test(hist$`ATRX`, hist$WT, alternative = "greater")
cac1_2_stat <- wilcox.test(hist$`cac-1_cac-2`, hist$WT, alternative = "greater")
cac1_suz12_stat <- wilcox.test(hist$`cac-1_suz-12`, hist$WT, alternative = "greater")
suz12_stat <- wilcox.test(hist$`suz-12`, hist$WT, alternative = "greater")
ash1_stat <- wilcox.test(hist$`ash-1`, hist$WT, alternative = "greater")

#Make a list of all wilcox results objects
wilcox_list <- list(set7_stat, cac1_stat, cac2_stat, cac3_stat, naf1_stat, naf2_stat, asf1_stat, atrx_stat, cac1_2_stat, cac1_suz12_stat, suz12_stat, ash1_stat)

#rename objects
names(wilcox_list) <- c("set7", "cac1", "cac2", "cac3", "naf1", "naf2", "asf1", "atrx", "cac-1_cac-2", "cac-1_suz-12", "suz-12", "ash-1")

#Extract relevant stats into a dataframe
wilcox_df <- do.call(rbind, lapply(wilcox_list, function(res) {
  data.frame(
    statistic = unname(res$statistic),
    p_value = res$p.value,
    alternative = res$alternative,
    method = res$method
  )
}))

# Set strain names as row names
rownames(wilcox_df) <- names(wilcox_list)

# Save to CSV
write.csv(wilcox_df, "Wilcoxon_summary_stats.csv", row.names = TRUE)
```

```{r Plotting data in Violin Plot, echo=FALSE}
# Define factor order and formatted x-axis labels
label_order <- c("WT", "set-7", "cac-1", "cac-2", "cac-3", "naf-1", "naf-2", "asf-1", "ATRX")
formatted_labels <- c(
  "WT",
  expression(italic("\u0394set-7")),
  expression(italic("\u0394cac-1")),
  expression(italic("\u0394cac-2")),
  expression(italic("\u0394cac-3")),
  expression(italic("\u0394naf-1")),
  expression(italic("\u0394naf-2")),
  expression(italic("\u0394asf-1")),
  expression(italic("\u0394atrx"))
)

#Group and summarize
total2 <- meltedAveragePRC2targetData %>%
  group_by(Sample)

total_dist <- meltedAveragePRC2targetData %>%
  group_by(Sample) %>%
  summarise(num = n())

#Set dodge for violin spacing
dodge <- position_dodge(width = 1)

#Build plot
violin <- total2 %>%
  left_join(total_dist) %>%
  arrange(factor(Sample, levels = label_order)) %>%
  mutate(Sample = factor(Sample, levels = label_order)) %>%
  ggplot(aes(x = Sample, y = Count)) +
  geom_violin(position = dodge, scale = "width", trim = FALSE) +
  stat_summary(fun = "mean", geom = "crossbar", width = 0.25, colour = "red") +
  scale_x_discrete(labels = formatted_labels) +
  labs(
    x = "Strain",
    y = expression("Expression Level (log"[2]~"(TPM+1))")) +
  theme_classic(base_size = 20) +
  theme(
    axis.text.x = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black")
  )

#check it
print(violin)

#save plot
ggsave("./CAF-1_K27_Violin_Paper_Final.pdf", plot=violin, width = 10, height = 8, unit="in",  dpi=400)
```

```{r Plotting data on heatmap, echo=FALSE}
##Keep Only Genes that are expressed in at least one sample (n = 564)
GenesWithChanges <- subset(AVERAGE_Prc2targetTPM, (rowSums(AVERAGE_Prc2targetTPM) > 0))

#plot in the desired column order; did this by subsetting the dataset based on sample list 'altorder' above
heatmap <- pheatmap::pheatmap(GenesWithChanges, color = colorRampPalette(rev(brewer.pal(n = 9, name ="RdBu")))(100), cellwidth = NA, cellheight = NA, scale = "row", cluster_rows = T, cluster_cols = F, clustering_method="centroid", clustering_distance_cols="euclidean", legend=T, show_rownames=F, show_colnames=T, fontsize_col=10, treeheight_row=0, treeheight_col=5, height = 1.5, width = 2.5, fontsize = 8)

#plot
heatmap_plot <- heatmap[[4]]

#save plot
ggsave(filename = "./CAF-1_K27_Heatmap_Paper_Final.pdf", plot = heatmap_plot, dpi=600, height=4, width=3)
```

```{r Plotting data in Violin Plot for Double Mutants, echo=FALSE}
# Define factor order and formatted x-axis labels
label_order <- c("WT", "suz-12", "cac-1", "cac-1_suz-12")
formatted_labels <- c(
  "WT",
  expression(italic("\u0394suz-12")),
  expression(italic("\u0394cac-1")),
  expression(italic("\u0394cac-1_suz-12"))
)

#Group and summarize
total2 <- meltedAverageDoublePRC2targetData %>%
  group_by(Sample)

total_dist <- meltedAverageDoublePRC2targetData %>%
  group_by(Sample) %>%
  summarise(num = n())

#Set dodge for violin spacing
dodge <- position_dodge(width = 1)

#Build plot
violin <- total2 %>%
  left_join(total_dist) %>%
  arrange(factor(Sample, levels = label_order)) %>%
  mutate(Sample = factor(Sample, levels = label_order)) %>%
  ggplot(aes(x = Sample, y = Count)) +
  geom_violin(position = dodge, scale = "width", trim = FALSE) +
  stat_summary(fun = "mean", geom = "crossbar", width = 0.25, colour = "red") +
  scale_x_discrete(labels = formatted_labels) +
  labs(
    x = "Strain",
    y = expression("Expression Level (log"[2]~"(TPM+1))")) +
  theme_classic(base_size = 20) +
  theme(
    axis.text.x = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black")
  )

#check it
print(violin)

#save plot
ggsave("./CAF-1_Double_Mutant_K27_Violin_Paper_Final.pdf", plot=violin, width = 10, height = 8, unit="in",  dpi=400)
```

```{r Plotting data on heatmap for Double Mutants, echo=FALSE}
##Keep Only Genes that are expressed in at least one sample (n = 501)
GenesWithChanges <- subset(AVERAGE_Prc2targetTPM_Double, (rowSums(AVERAGE_Prc2targetTPM_Double) > 0))

#plot in the desired column order; did this by subsetting the dataset based on sample list 'altorder' above
heatmap <- pheatmap::pheatmap(GenesWithChanges, color = colorRampPalette(rev(brewer.pal(n = 9, name ="RdBu")))(100), cellwidth = NA, cellheight = NA, scale = "row", cluster_rows = T, cluster_cols = F, clustering_method="centroid", clustering_distance_cols="euclidean", legend=T, show_rownames=F, show_colnames=T, fontsize_col=10, treeheight_row=0, treeheight_col=5, height = 1.5, width = 2.5, fontsize = 8)

#plot
heatmap_plot <- heatmap[[4]]

#save plot
ggsave(filename = "./CAF-1_Double_Mutant_K27_Heatmap_Paper_Final.pdf", plot = heatmap_plot, dpi=600, height=4, width=3)
```

#Moving on to DESeq2
##We will use this to determine the genes significantly DE (p < 0.05 & < 0.01)
##These lists will be used to make volcano plots, Venn Diagrams, and for GO Analysis

```{r importing counts}
#Note: file was already read in, using a different variable name for DESeq to avoid confusion this was already done
## change ncol to match number of columns
cts_numeric <- matrix(as.numeric(Interactors_countsOnly), ncol = 42)
dimnames(cts_numeric) <- list(rownames(Interactors_countsOnly), colnames(Interactors_countsOnly))

#Rename columns
colnames(cts_numeric) <- names[,1]
```

```{r group samples, echo=FALSE}
#grouping samples by strain for DESeq analysis
##Note, I am leaving old cac-1 strain in analysis since the counts are in the file, but will not be utilizing the data for figures. I will be using the "cac-1_new" data instead
cac1 <- rowMeans(cts_numeric[, 29:31], na.rm = TRUE)
cac2 <- rowMeans(cts_numeric[, 4:6], na.rm = TRUE)
cac3 <- rowMeans(cts_numeric[, 7:9], na.rm = TRUE)
naf1 <- rowMeans(cts_numeric[, 10:12], na.rm = TRUE)
asf1 <- rowMeans(cts_numeric[, 13:15], na.rm = TRUE)
naf2 <- rowMeans(cts_numeric[, 16:18], na.rm = TRUE)
ATRX <- rowMeans(cts_numeric[, 19:21], na.rm = TRUE)
WT <- rowMeans(cts_numeric[, 22:25], na.rm = TRUE)
set7 <- rowMeans(cts_numeric[, 26:28], na.rm = TRUE)
cac1_cac2 <- rowMeans(cts_numeric[, 32:34], na.rm = TRUE)
cac1_suz12 <- rowMeans(cts_numeric[, 35:37], na.rm = TRUE)
suz12 <- rowMeans(cts_numeric[, 38:40], na.rm = TRUE)
ash1 <- rowMeans(cts_numeric[, 41:42], na.rm = TRUE)
```

```{r load coldata, echo = FALSE}
#Loading in column data file. This will be used to assign the counts samples to the correct strains in DESeq
##Make sure you have the desired strain listed as a control in this sheet
coldata <- read.csv("../coldata.csv", header= TRUE, row.names = 1)
coldata$condition <- factor(coldata$condition)
coldata$type <- factor(coldata$type)

#Confirm that the number and order of samples are the same in each dataset before proceeding. Important, because this will throw off/fail DESeq otherwise
all(rownames(coldata) %in% colnames(cts_numeric))
all(rownames(coldata) == colnames(cts_numeric))

```

```{r, Run DESeq analysis, echo=FALSE}
# --- your setup code (unchanged) ---
cts_numeric_pseudo = cts_numeric + 1
dds <- DESeqDataSetFromMatrix(countData = round(cts_numeric_pseudo),
                              colData = coldata,
                              design = ~ condition)
keep <- rowSums(counts(dds)) > 38
dds <- dds[keep,]
dds$condition <- relevel(dds$condition, ref = "WT")

dds2 <- DESeq(dds)
plotDispEsts(dds2)

```

```{r create csvs with DE genes for each strain, echo=FALSE}
export_deseq_csvs <- function(
  dds2,
  ref_level = "WT",
  strains,
  neurospora_bed_path,
  prc2_targets_path,
  lfc_cutoff = 1.0,
  alphas = c(0.05, 0.01)
) {
  ## ---- load mappings ----
  neurospora  <- read.table(neurospora_bed_path, header = TRUE, sep = "\t",
                            check.names = FALSE, stringsAsFactors = FALSE)
  Prc2targets <- read.table(prc2_targets_path,   header = FALSE, sep = "\t",
                            check.names = FALSE, stringsAsFactors = FALSE)

  choose_first <- function(nms, opts) opts[opts %in% nms][1]
  idcol_bed <- choose_first(names(neurospora),  c("VALUE","gene_id","NCU","Attributes","V11","V4"))
  idcol_prc <- choose_first(names(Prc2targets), c("V11","VALUE","gene_id","NCU","Attributes","V4"))
  if (is.na(idcol_bed) || is.na(idcol_prc)) stop("Could not find ID columns in BED/PRC2 files.")
  prc2_ids <- unique(Prc2targets[[idcol_prc]])

  ## ---- helpers ----
  alpha_lab  <- function(a) formatC(a, format = "f", digits = 2)
  lfc_tag    <- function(x) paste0("ge_", gsub("\\.", "p", sprintf("%.1f", x)))  # e.g., ge_1p0
  sanitize   <- function(x) gsub("[^A-Za-z0-9._-]+", "_", x)
  safe_write <- function(df, fname) utils::write.csv(df, file = sanitize(fname), row.names = FALSE)
  bed_from_ids <- function(ids) neurospora[neurospora[[idcol_bed]] %in% unique(ids), , drop = FALSE]

  # RAW (unshrunken) DESeq2 results used here
  res_for <- function(strain, alpha) {
    df <- as.data.frame(DESeq2::results(dds2, contrast = c("condition", strain, ref_level), alpha = alpha))
    df$NCU <- rownames(df)
    df <- df[!is.na(df$padj), ]
    sig <- df[df$padj < alpha & abs(df$log2FoldChange) >= lfc_cutoff, ]
    up  <- sig[sig$log2FoldChange >=  lfc_cutoff, ]
    dn  <- sig[sig$log2FoldChange <= -lfc_cutoff, ]
    list(all = sig, up = up, dn = dn, full = df)
  }

  write_deg_csvs <- function(strain, alpha, reslist) {
    a   <- alpha_lab(alpha)
    lt  <- lfc_tag(lfc_cutoff)
    base <- paste0(strain, "__alpha", a)

    ## 1) ALL results (no cutoffs), RAW LFCs
    res_all <- as.data.frame(DESeq2::results(dds2, contrast = c("condition", strain, ref_level),
                                             independentFiltering = FALSE))
    res_all$NCU <- rownames(res_all)
    safe_write(res_all[, c("NCU", setdiff(names(res_all), "NCU"))],
               sprintf("%s__ALL_results_LFC.csv", base))

    ## 2) DE table: padj<alpha & |LFC|>=cutoff (both directions)
    safe_write(reslist$all[, c("NCU", setdiff(names(reslist$all), "NCU"))],
               sprintf("%s__DE_absLFC_%s.csv", base, lt))

    ## 3) UP results (actual rows meeting cutoffs)
    if (nrow(reslist$up)) {
      safe_write(reslist$up[, c("NCU", setdiff(names(reslist$up), "NCU"))],
                 sprintf("%s__UP_results_LFC_%s.csv", base, lt))
    } else {
      safe_write(data.frame(), sprintf("%s__UP_results_LFC_%s.csv", base, lt))
    }

    ## 4) DOWN results (actual rows meeting cutoffs)
    if (nrow(reslist$dn)) {
      safe_write(reslist$dn[, c("NCU", setdiff(names(reslist$dn), "NCU"))],
                 sprintf("%s__DOWN_results_LFC_%s.csv", base, lt))
    } else {
      safe_write(data.frame(), sprintf("%s__DOWN_results_LFC_%s.csv", base, lt))
    }

    ## 5) UP gene list (BED-mapped rows)
    up_ids <- unique(reslist$up$NCU)
    safe_write(bed_from_ids(up_ids), sprintf("%s__UP_genes.csv", base))

    ## 6) DOWN gene list (BED-mapped rows)
    dn_ids <- unique(reslist$dn$NCU)
    safe_write(bed_from_ids(dn_ids), sprintf("%s__DOWN_genes.csv", base))

    ## 7) K27-UP gene list (UP ∩ PRC2 targets)
    safe_write(bed_from_ids(intersect(up_ids, prc2_ids)), sprintf("%s__K27_UP_genes.csv", base))

    ## 8) nonK27-UP gene list (UP \ PRC2 targets)
    nonk27_up_ids <- setdiff(up_ids, prc2_ids)
    safe_write(bed_from_ids(nonk27_up_ids), sprintf("%s__nonK27_UP_genes.csv", base))

    invisible(list(up_ids = up_ids, dn_ids = dn_ids, de_ids = unique(reslist$all$NCU)))
  }

  ## ---- main per-strain loop ----
  available_levels <- levels(colData(dds2)$condition)
  deg_map <- list()
  for (alpha in alphas) {
    a <- alpha_lab(alpha)
    for (s in strains) {
      if (!(s %in% available_levels)) {
        warning("Skipping '", s, "' — not found in colData(dds2)$condition.")
        next
      }
      rl <- res_for(s, alpha)
      id_sets <- write_deg_csvs(s, alpha, rl)
      deg_map[[paste(s, a, sep = "|")]] <- id_sets
    }
  }

  ## ---- CAF-1 combos: pairwise (cac1_cac2) & triple (cac1_cac2_cac3) ----
  # Now includes UP and DOWN intersections (plus existing DE and nonK27_UP)
  caf1 <- c("cac1","cac2","cac3")
  have_caf1 <- all(caf1 %in% strains) && all(caf1 %in% available_levels)
  if (have_caf1) {
    for (alpha in alphas) {
      a <- alpha_lab(alpha)
      key <- function(s) paste(s, a, sep = "|")

      up_sets <- lapply(caf1, function(s) deg_map[[key(s)]]$up_ids)
      dn_sets <- lapply(caf1, function(s) deg_map[[key(s)]]$dn_ids)
      de_sets <- lapply(caf1, function(s) deg_map[[key(s)]]$de_ids)

      ## pairwise: cac1 & cac2 (UP)
      safe_write(bed_from_ids(intersect(up_sets[[1]], up_sets[[2]])),
                 sprintf("cac1_cac2__alpha%s__UP_intersection.csv", a))
      ## pairwise: cac1 & cac2 (DOWN)
      safe_write(bed_from_ids(intersect(dn_sets[[1]], dn_sets[[2]])),
                 sprintf("cac1_cac2__alpha%s__DOWN_intersection.csv", a))

      ## pairwise: cac1 & cac2 (DE)
      safe_write(bed_from_ids(intersect(de_sets[[1]], de_sets[[2]])),
                 sprintf("cac1_cac2__alpha%s__DE_intersection.csv", a))

      ## pairwise: cac1 & cac2 (nonK27_UP)
      safe_write(bed_from_ids(setdiff(intersect(up_sets[[1]], up_sets[[2]]), prc2_ids)),
                 sprintf("cac1_cac2__alpha%s__nonK27_UP_intersection.csv", a))

      ## triple: cac1 & cac2 & cac3 (UP)
      safe_write(bed_from_ids(Reduce(intersect, up_sets)),
                 sprintf("cac1_cac2_cac3__alpha%s__UP_intersection.csv", a))
      ## triple: cac1 & cac2 & cac3 (DOWN)
      safe_write(bed_from_ids(Reduce(intersect, dn_sets)),
                 sprintf("cac1_cac2_cac3__alpha%s__DOWN_intersection.csv", a))

      ## triple: cac1 & cac2 & cac3 (DE)
      safe_write(bed_from_ids(Reduce(intersect, de_sets)),
                 sprintf("cac1_cac2_cac3__alpha%s__DE_intersection.csv", a))

      ## triple: cac1 & cac2 & cac3 (nonK27_UP)
      safe_write(bed_from_ids(setdiff(Reduce(intersect, up_sets), prc2_ids)),
                 sprintf("cac1_cac2_cac3__alpha%s__nonK27_UP_intersection.csv", a))
    }
  } else {
    message("CAF-1 combo outputs skipped (one or more of cac1/cac2/cac3 missing).")
  }

  invisible(deg_map)
}

# call stays the same
deg_map <- export_deseq_csvs(
  dds2,
  ref_level = "WT",
  strains = c("cac1","cac2","cac3","set7","naf1","naf2","asf1","atrx","cac1_cac2","cac1_suz12","suz12","ash1"),
  neurospora_bed_path = "../bed_files/neurospora_genes_edit.bed",
  prc2_targets_path   = "../bed_files/H3K27me3_methylated_genes_FINAL.bed",
  lfc_cutoff = 1.0,
  alphas = c(0.05, 0.01)
)

```

#GO Analysis
```{r, my script}
# GO BP enrichment + lollipop plot for Δcac-1 & Δcac-2 DE genes
# Requires:
#   - Nc_GO_mapping.txt  (your FungiDB export)
#   - cac1_cac2__alpha0.05__DE_intersection.csv  (shared DE genes)

## --- packages ---
# cran_pkgs <- c("readr","dplyr","tidyr","ggplot2")
# miss <- setdiff(cran_pkgs, rownames(installed.packages()))
# if (length(miss)) install.packages(miss, repos = "https://cloud.r-project.org")
# if (!requireNamespace("BiocManager", quietly = TRUE))
#   install.packages("BiocManager", repos = "https://cloud.r-project.org")
# BiocManager::install(c("clusterProfiler","GO.db","AnnotationDbi"), ask = FALSE, update = FALSE)
# 
# suppressPackageStartupMessages({
#   library(readr); library(dplyr); library(tidyr); library(ggplot2)
#   library(clusterProfiler); library(GO.db); library(AnnotationDbi)
# })

## --- paths / outputs ---
fungidb_go_path <- "Nc_GO_mapping.txt"
de_csv_path     <- "../CAF-1_RNA-seq_Analysis/csv_files/cac1_cac2__alpha0.05__DE_intersection.csv"
out_tag  <- "CAF1_shared_GO"
out_res  <- paste0(out_tag, "_BP_enrichment.tsv")
out_plot <- paste0(out_tag, "_BP_lollipop.pdf")
top_n_terms <- 20

## --- read GO mapping (FungiDB tab with separate ID/name cols) ---
stopifnot(file.exists(fungidb_go_path))
go_raw <- readr::read_tsv(fungidb_go_path, show_col_types = FALSE)

gene_col <- intersect(names(go_raw), c("Gene ID","gene_id","Genes","ID"))[1]
if (is.na(gene_col)) stop("Could not find a 'Gene ID' column in Nc_GO_mapping.txt")

pairs <- list(
  list(id="Curated GO Process IDs",   nm="Curated GO Processes",   ns="BP"),
  list(id="Computed GO Process IDs",  nm="Computed GO Processes",   ns="BP"),
  list(id="Curated GO Component IDs", nm="Curated GO Components",   ns="CC"),
  list(id="Computed GO Component IDs",nm="Computed GO Components",  ns="CC"),
  list(id="Curated GO Function IDs",  nm="Curated GO Functions",    ns="MF"),
  list(id="Computed GO Function IDs", nm="Computed GO Functions",   ns="MF")
)

make_long <- function(idcol, nmcol, ns){
  if (!(idcol %in% names(go_raw))) return(NULL)
  tmp <- go_raw %>%
    dplyr::transmute(
      gene_id = .data[[gene_col]],
      go_id   = as.character(.data[[idcol]]),
      go_name = if (nmcol %in% names(go_raw)) as.character(.data[[nmcol]]) else NA_character_
    ) %>%
    dplyr::filter(!is.na(go_id), nzchar(go_id)) %>%
    tidyr::separate_rows(go_id,  sep = "\\s*;\\s*") %>%
    tidyr::separate_rows(go_name, sep = "\\s*;\\s*") %>%
    dplyr::mutate(namespace = ns) %>%
    dplyr::distinct()
  tmp
}

go_map <- do.call(
  dplyr::bind_rows,
  lapply(pairs, function(p) make_long(p$id, p$nm, p$ns))
) %>%
  dplyr::filter(grepl("^GO:", go_id)) %>%
  dplyr::distinct()

# fill missing names from GO.db
need_names <- setdiff(unique(go_map$go_id), unique(go_map$go_id[!is.na(go_map$go_name) & nzchar(go_map$go_name)]))
if (length(need_names)) {
  dbn <- AnnotationDbi::select(GO.db, keys = need_names, keytype = "GOID", columns = "TERM")
  dbn <- dplyr::rename(dbn, go_id = GOID, go_name_db = TERM)
  go_map <- go_map %>%
    dplyr::left_join(dbn, by = "go_id") %>%
    dplyr::mutate(go_name = dplyr::coalesce(go_name, go_name_db)) %>%
    dplyr::select(-go_name_db)
}

## --- read shared DE genes, auto-detect ID column ---
stopifnot(file.exists(de_csv_path))
de_df <- readr::read_csv(de_csv_path, show_col_types = FALSE)

guess_id_col <- function(df){
  # prefer obvious columns
  cand <- intersect(names(df), c("gene_id","Gene ID","ID","Genes","NCU","NCU_ID","NCUid","Attributes","VALUE"))
  if (length(cand)) return(cand[1])
  # else pick column with most NCU-like entries
  score <- sapply(df, function(x) sum(grepl("^NCU\\d{5}$", as.character(x))))
  names(which.max(score))
}
de_col <- guess_id_col(de_df)
genes_shared <- unique(as.character(de_df[[de_col]]))

## --- universe: all genes that have any GO term (quick + conservative) ---
universe_genes <- unique(as.character(go_map$gene_id))

## --- TERM2GENE / TERM2NAME for Biological Process only ---
TERM2GENE <- go_map %>%
  dplyr::filter(tolower(namespace) %in% c("bp","biological process","biological_process","p")) %>%
  dplyr::select(go_id, gene_id) %>% dplyr::distinct()
TERM2NAME <- go_map %>%
  dplyr::select(go_id, go_name) %>% dplyr::distinct()

## --- enrichment (BH FDR; we’ll filter to FDR <= 0.05 for plotting) ---
enr_bp <- clusterProfiler::enricher(
  gene          = genes_shared,
  universe      = universe_genes,
  TERM2GENE     = TERM2GENE,
  TERM2NAME     = TERM2NAME,
  pAdjustMethod = "BH",
  pvalueCutoff  = 1.0,
  qvalueCutoff  = 1.0
)
if (is.null(enr_bp) || nrow(enr_bp@result) == 0)
  stop("No BP terms returned. Check IDs vs GO mapping.")

## --- compute fold-enrichment and tidy ---
res <- enr_bp@result %>%
  dplyr::mutate(
    GeneRatioNum = sapply(strsplit(GeneRatio, "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    BgRatioNum   = sapply(strsplit(BgRatio,   "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    FoldEnrich   = GeneRatioNum / BgRatioNum,
    FDR          = p.adjust(pvalue, "BH"),
    minusLog10FDR = -log10(FDR),
    Term = dplyr::coalesce(Description, ID)   # Description is the term name from TERM2NAME
  ) %>%
  dplyr::arrange(dplyr::desc(FoldEnrich))

readr::write_tsv(res, out_res)

## --- lollipop plot: names on y, bigger size spread, vivid colors; FDR-filtered ---
max_chars <- 37
shorten <- function(x, n = max_chars) ifelse(nchar(x) > n, paste0(substr(x, 1, n - 1), "…"), x)

plot_df <- res %>%
  dplyr::filter(FDR <= 0.05) %>%                # significant terms only
  dplyr::slice_head(n = top_n_terms) %>%
  dplyr::mutate(
    LabelShow = shorten(Term),                  # what you SEE (no GO IDs)
    LabelKey  = paste0(Term, "\u200B", dplyr::row_number()), # invisible uniqueness
    Label     = factor(LabelKey, levels = rev(LabelKey))
  )

lab_map <- setNames(plot_df$LabelShow, plot_df$LabelKey)

p <- ggplot(plot_df, aes(FoldEnrich, Label)) +
  geom_segment(aes(x = 0, xend = FoldEnrich, yend = Label), linewidth = 0.8) +
  geom_point(aes(size = Count, color = minusLog10FDR)) +
  scale_size(range = c(3, 10), name = "N. of Genes") +
  scale_color_viridis_c(option = "plasma", name = expression(-log[10]("FDR"))) +
  scale_y_discrete(labels = lab_map) +          # show shortened names only
  labs(x = "Fold Enrichment", y = NULL,
       title = expression("GO Biological Process Enrichment (cac-1 cac-2 DE)")) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y  = element_text(size = 6),
    axis.text.x  = element_text(size = 9),
    legend.text  = element_text(size = 9),
    legend.title = element_text(size = 10),
    plot.title   = element_text(size = 12, hjust = 0.5),
    panel.grid.major.y = element_blank()
  )

ggplot2::ggsave(out_plot, p, width = 7, height = 5)
p

########################################################################################
###Splitting UP + Downregulated genes in CAF-1 mutants

## --- paths / outputs ---
fungidb_go_path <- "Nc_GO_mapping.txt"
de_csv_path <- "../CAF-1_RNA-seq_Analysis/csv_files/cac1_cac2__alpha0.05__UP_intersection.csv"
out_tag <- "CAF1_shared_UP_GO"
out_res <- paste0(out_tag, "_BP_enrichment.tsv")
out_plot <- paste0(out_tag, "_BP_lollipop.pdf")
top_n_terms <- 20

## --- read GO mapping (FungiDB tab with separate ID/name cols) ---
stopifnot(file.exists(fungidb_go_path))
go_raw <- readr::read_tsv(fungidb_go_path, show_col_types = FALSE)

gene_col <- intersect(names(go_raw), c("Gene ID","gene_id","Genes","ID"))[1]
if (is.na(gene_col)) stop("Could not find a 'Gene ID' column in Nc_GO_mapping.txt")

pairs <- list(
  list(id="Curated GO Process IDs",   nm="Curated GO Processes",   ns="BP"),
  list(id="Computed GO Process IDs",  nm="Computed GO Processes",   ns="BP"),
  list(id="Curated GO Component IDs", nm="Curated GO Components",   ns="CC"),
  list(id="Computed GO Component IDs",nm="Computed GO Components",  ns="CC"),
  list(id="Curated GO Function IDs",  nm="Curated GO Functions",    ns="MF"),
  list(id="Computed GO Function IDs", nm="Computed GO Functions",   ns="MF")
)

make_long <- function(idcol, nmcol, ns){
  if (!(idcol %in% names(go_raw))) return(NULL)
  tmp <- go_raw %>%
    dplyr::transmute(
      gene_id = .data[[gene_col]],
      go_id   = as.character(.data[[idcol]]),
      go_name = if (nmcol %in% names(go_raw)) as.character(.data[[nmcol]]) else NA_character_
    ) %>%
    dplyr::filter(!is.na(go_id), nzchar(go_id)) %>%
    tidyr::separate_rows(go_id,  sep = "\\s*;\\s*") %>%
    tidyr::separate_rows(go_name, sep = "\\s*;\\s*") %>%
    dplyr::mutate(namespace = ns) %>%
    dplyr::distinct()
  tmp
}

go_map <- do.call(
  dplyr::bind_rows,
  lapply(pairs, function(p) make_long(p$id, p$nm, p$ns))
) %>%
  dplyr::filter(grepl("^GO:", go_id)) %>%
  dplyr::distinct()

# fill missing names from GO.db
need_names <- setdiff(unique(go_map$go_id), unique(go_map$go_id[!is.na(go_map$go_name) & nzchar(go_map$go_name)]))
if (length(need_names)) {
  dbn <- AnnotationDbi::select(GO.db, keys = need_names, keytype = "GOID", columns = "TERM")
  dbn <- dplyr::rename(dbn, go_id = GOID, go_name_db = TERM)
  go_map <- go_map %>%
    dplyr::left_join(dbn, by = "go_id") %>%
    dplyr::mutate(go_name = dplyr::coalesce(go_name, go_name_db)) %>%
    dplyr::select(-go_name_db)
}

## --- read shared DE genes, auto-detect ID column ---
stopifnot(file.exists(de_csv_path))
de_df <- readr::read_csv(de_csv_path, show_col_types = FALSE)

guess_id_col <- function(df){
  # prefer obvious columns
  cand <- intersect(names(df), c("gene_id","Gene ID","ID","Genes","NCU","NCU_ID","NCUid","Attributes","VALUE"))
  if (length(cand)) return(cand[1])
  # else pick column with most NCU-like entries
  score <- sapply(df, function(x) sum(grepl("^NCU\\d{5}$", as.character(x))))
  names(which.max(score))
}
de_col <- guess_id_col(de_df)
genes_shared <- unique(as.character(de_df[[de_col]]))

## --- universe: all genes that have any GO term (quick + conservative) ---
universe_genes <- unique(as.character(go_map$gene_id))

## --- TERM2GENE / TERM2NAME for Biological Process only ---
TERM2GENE <- go_map %>%
  dplyr::filter(tolower(namespace) %in% c("bp","biological process","biological_process","p")) %>%
  dplyr::select(go_id, gene_id) %>% dplyr::distinct()
TERM2NAME <- go_map %>%
  dplyr::select(go_id, go_name) %>% dplyr::distinct()

## --- enrichment (BH FDR; we’ll filter to FDR <= 0.05 for plotting) ---
enr_bp <- clusterProfiler::enricher(
  gene          = genes_shared,
  universe      = universe_genes,
  TERM2GENE     = TERM2GENE,
  TERM2NAME     = TERM2NAME,
  pAdjustMethod = "BH",
  pvalueCutoff  = 1.0,
  qvalueCutoff  = 1.0
)
if (is.null(enr_bp) || nrow(enr_bp@result) == 0)
  stop("No BP terms returned. Check IDs vs GO mapping.")

## --- compute fold-enrichment and tidy ---
res <- enr_bp@result %>%
  dplyr::mutate(
    GeneRatioNum = sapply(strsplit(GeneRatio, "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    BgRatioNum   = sapply(strsplit(BgRatio,   "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    FoldEnrich   = GeneRatioNum / BgRatioNum,
    FDR          = p.adjust(pvalue, "BH"),
    minusLog10FDR = -log10(FDR),
    Term = dplyr::coalesce(Description, ID)   # Description is the term name from TERM2NAME
  ) %>%
  dplyr::arrange(dplyr::desc(FoldEnrich))

readr::write_tsv(res, out_res)

## --- lollipop plot: names on y, bigger size spread, vivid colors; FDR-filtered ---
max_chars <- 37
shorten <- function(x, n = max_chars) ifelse(nchar(x) > n, paste0(substr(x, 1, n - 1), "…"), x)

plot_df <- res %>%
  dplyr::filter(FDR <= 0.05) %>%                # significant terms only
  dplyr::slice_head(n = top_n_terms) %>%
  dplyr::mutate(
    LabelShow = shorten(Term),                  # what you SEE (no GO IDs)
    LabelKey  = paste0(Term, "\u200B", dplyr::row_number()), # invisible uniqueness
    Label     = factor(LabelKey, levels = rev(LabelKey))
  )

lab_map <- setNames(plot_df$LabelShow, plot_df$LabelKey)

p <- ggplot(plot_df, aes(FoldEnrich, Label)) +
  geom_segment(aes(x = 0, xend = FoldEnrich, yend = Label), linewidth = 0.8) +
  geom_point(aes(size = Count, color = minusLog10FDR)) +
  scale_size(range = c(3, 10), name = "N. of Genes") +
  scale_color_viridis_c(option = "plasma", name = expression(-log[10]("FDR"))) +
  scale_y_discrete(labels = lab_map) +          # show shortened names only
  labs(x = "Fold Enrichment", y = NULL,
       title = expression("GO Biological Process Enrichment (cac-1 cac-2 UP)")) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y  = element_text(size = 6),
    axis.text.x  = element_text(size = 9),
    legend.text  = element_text(size = 9),
    legend.title = element_text(size = 10),
    plot.title   = element_text(size = 12, hjust = 0.5),
    panel.grid.major.y = element_blank()
  )

ggplot2::ggsave(out_plot, p, width = 7, height = 5)
p

#######

## --- paths / outputs ---
fungidb_go_path <- "Nc_GO_mapping.txt"
de_csv_path     <- "../CAF-1_RNA-seq_Analysis/csv_files/cac1_cac2__alpha0.05__DOWN_intersection.csv"
out_tag  <- "CAF1_shared_DOWN_GO"
out_res  <- paste0(out_tag, "_BP_enrichment.tsv")
out_plot <- paste0(out_tag, "_BP_lollipop.pdf")
top_n_terms <- 20

## --- read GO mapping (FungiDB tab with separate ID/name cols) ---
stopifnot(file.exists(fungidb_go_path))
go_raw <- readr::read_tsv(fungidb_go_path, show_col_types = FALSE)

gene_col <- intersect(names(go_raw), c("Gene ID","gene_id","Genes","ID"))[1]
if (is.na(gene_col)) stop("Could not find a 'Gene ID' column in Nc_GO_mapping.txt")

pairs <- list(
  list(id="Curated GO Process IDs",   nm="Curated GO Processes",   ns="BP"),
  list(id="Computed GO Process IDs",  nm="Computed GO Processes",   ns="BP"),
  list(id="Curated GO Component IDs", nm="Curated GO Components",   ns="CC"),
  list(id="Computed GO Component IDs",nm="Computed GO Components",  ns="CC"),
  list(id="Curated GO Function IDs",  nm="Curated GO Functions",    ns="MF"),
  list(id="Computed GO Function IDs", nm="Computed GO Functions",   ns="MF")
)

make_long <- function(idcol, nmcol, ns){
  if (!(idcol %in% names(go_raw))) return(NULL)
  tmp <- go_raw %>%
    dplyr::transmute(
      gene_id = .data[[gene_col]],
      go_id   = as.character(.data[[idcol]]),
      go_name = if (nmcol %in% names(go_raw)) as.character(.data[[nmcol]]) else NA_character_
    ) %>%
    dplyr::filter(!is.na(go_id), nzchar(go_id)) %>%
    tidyr::separate_rows(go_id,  sep = "\\s*;\\s*") %>%
    tidyr::separate_rows(go_name, sep = "\\s*;\\s*") %>%
    dplyr::mutate(namespace = ns) %>%
    dplyr::distinct()
  tmp
}

go_map <- do.call(
  dplyr::bind_rows,
  lapply(pairs, function(p) make_long(p$id, p$nm, p$ns))
) %>%
  dplyr::filter(grepl("^GO:", go_id)) %>%
  dplyr::distinct()

# fill missing names from GO.db
need_names <- setdiff(unique(go_map$go_id), unique(go_map$go_id[!is.na(go_map$go_name) & nzchar(go_map$go_name)]))
if (length(need_names)) {
  dbn <- AnnotationDbi::select(GO.db, keys = need_names, keytype = "GOID", columns = "TERM")
  dbn <- dplyr::rename(dbn, go_id = GOID, go_name_db = TERM)
  go_map <- go_map %>%
    dplyr::left_join(dbn, by = "go_id") %>%
    dplyr::mutate(go_name = dplyr::coalesce(go_name, go_name_db)) %>%
    dplyr::select(-go_name_db)
}

## --- read shared DE genes, auto-detect ID column ---
stopifnot(file.exists(de_csv_path))
de_df <- readr::read_csv(de_csv_path, show_col_types = FALSE)

guess_id_col <- function(df){
  # prefer obvious columns
  cand <- intersect(names(df), c("gene_id","Gene ID","ID","Genes","NCU","NCU_ID","NCUid","Attributes","VALUE"))
  if (length(cand)) return(cand[1])
  # else pick column with most NCU-like entries
  score <- sapply(df, function(x) sum(grepl("^NCU\\d{5}$", as.character(x))))
  names(which.max(score))
}
de_col <- guess_id_col(de_df)
genes_shared <- unique(as.character(de_df[[de_col]]))

## --- universe: all genes that have any GO term (quick + conservative) ---
universe_genes <- unique(as.character(go_map$gene_id))

## --- TERM2GENE / TERM2NAME for Biological Process only ---
TERM2GENE <- go_map %>%
  dplyr::filter(tolower(namespace) %in% c("bp","biological process","biological_process","p")) %>%
  dplyr::select(go_id, gene_id) %>% dplyr::distinct()
TERM2NAME <- go_map %>%
  dplyr::select(go_id, go_name) %>% dplyr::distinct()

## --- enrichment (BH FDR; we’ll filter to FDR <= 0.05 for plotting) ---
enr_bp <- clusterProfiler::enricher(
  gene          = genes_shared,
  universe      = universe_genes,
  TERM2GENE     = TERM2GENE,
  TERM2NAME     = TERM2NAME,
  pAdjustMethod = "BH",
  pvalueCutoff  = 1.0,
  qvalueCutoff  = 1.0
)
if (is.null(enr_bp) || nrow(enr_bp@result) == 0)
  stop("No BP terms returned. Check IDs vs GO mapping.")

## --- compute fold-enrichment and tidy ---
res <- enr_bp@result %>%
  dplyr::mutate(
    GeneRatioNum = sapply(strsplit(GeneRatio, "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    BgRatioNum   = sapply(strsplit(BgRatio,   "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    FoldEnrich   = GeneRatioNum / BgRatioNum,
    FDR          = p.adjust(pvalue, "BH"),
    minusLog10FDR = -log10(FDR),
    Term = dplyr::coalesce(Description, ID)   # Description is the term name from TERM2NAME
  ) %>%
  dplyr::arrange(dplyr::desc(FoldEnrich))

readr::write_tsv(res, out_res)

## --- lollipop plot: names on y, bigger size spread, vivid colors; FDR-filtered ---
max_chars <- 37
shorten <- function(x, n = max_chars) ifelse(nchar(x) > n, paste0(substr(x, 1, n - 1), "…"), x)

plot_df <- res %>%
  dplyr::filter(FDR <= 0.05) %>%                # significant terms only
  dplyr::slice_head(n = top_n_terms) %>%
  dplyr::mutate(
    LabelShow = shorten(Term),                  # what you SEE (no GO IDs)
    LabelKey  = paste0(Term, "\u200B", dplyr::row_number()), # invisible uniqueness
    Label     = factor(LabelKey, levels = rev(LabelKey))
  )

lab_map <- setNames(plot_df$LabelShow, plot_df$LabelKey)

p <- ggplot(plot_df, aes(FoldEnrich, Label)) +
  geom_segment(aes(x = 0, xend = FoldEnrich, yend = Label), linewidth = 0.8) +
  geom_point(aes(size = Count, color = minusLog10FDR)) +
  scale_size(range = c(3, 10), name = "N. of Genes") +
  scale_color_viridis_c(option = "plasma", name = expression(-log[10]("FDR"))) +
  scale_y_discrete(labels = lab_map) +          # show shortened names only
  labs(x = "Fold Enrichment", y = NULL,
       title = expression("GO Biological Process Enrichment (cac-1 cac-2 DOWN)")) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y  = element_text(size = 6),
    axis.text.x  = element_text(size = 9),
    legend.text  = element_text(size = 9),
    legend.title = element_text(size = 10),
    plot.title   = element_text(size = 12, hjust = 0.5),
    panel.grid.major.y = element_blank()
  )

ggplot2::ggsave(out_plot, p, width = 7, height = 5)
p

#########################################################################################

# GO BP enrichment + lollipop plot for Δcac-1Δsuz-12 UP genes
# Requires:
#   - Nc_GO_mapping.txt
#   - cac1_suz12__alpha0.05__UP_genes.csv

## --- paths / outputs ---
fungidb_go_path <- "Nc_GO_mapping.txt"
de_csv_path     <- "../CAF-1_RNA-seq_Analysis/csv_files/cac1_suz12__alpha0.05__UP_genes.csv"
out_tag  <- "cac1_suz12_UP_GO"
out_res  <- paste0(out_tag, "_BP_enrichment.tsv")
out_plot <- paste0(out_tag, "_BP_lollipop.pdf")
top_n_terms <- 20

## --- read GO mapping (FungiDB tab with separate ID/name cols) ---
stopifnot(file.exists(fungidb_go_path))
go_raw <- readr::read_tsv(fungidb_go_path, show_col_types = FALSE)

gene_col <- intersect(names(go_raw), c("Gene ID","gene_id","Genes","ID"))[1]
if (is.na(gene_col)) stop("Could not find a 'Gene ID' column in Nc_GO_mapping.txt")

pairs <- list(
  list(id="Curated GO Process IDs",   nm="Curated GO Processes",   ns="BP"),
  list(id="Computed GO Process IDs",  nm="Computed GO Processes",   ns="BP"),
  list(id="Curated GO Component IDs", nm="Curated GO Components",   ns="CC"),
  list(id="Computed GO Component IDs",nm="Computed GO Components",  ns="CC"),
  list(id="Curated GO Function IDs",  nm="Curated GO Functions",    ns="MF"),
  list(id="Computed GO Function IDs", nm="Computed GO Functions",   ns="MF")
)

make_long <- function(idcol, nmcol, ns){
  if (!(idcol %in% names(go_raw))) return(NULL)
  tmp <- go_raw %>%
    dplyr::transmute(
      gene_id = .data[[gene_col]],
      go_id   = as.character(.data[[idcol]]),
      go_name = if (nmcol %in% names(go_raw)) as.character(.data[[nmcol]]) else NA_character_
    ) %>%
    dplyr::filter(!is.na(go_id), nzchar(go_id)) %>%
    tidyr::separate_rows(go_id,  sep = "\\s*;\\s*") %>%
    tidyr::separate_rows(go_name, sep = "\\s*;\\s*") %>%
    dplyr::mutate(namespace = ns) %>%
    dplyr::distinct()
  tmp
}

go_map <- do.call(
  dplyr::bind_rows,
  lapply(pairs, function(p) make_long(p$id, p$nm, p$ns))
) %>%
  dplyr::filter(grepl("^GO:", go_id)) %>%
  dplyr::distinct()

# fill missing names from GO.db
need_names <- setdiff(unique(go_map$go_id), unique(go_map$go_id[!is.na(go_map$go_name) & nzchar(go_map$go_name)]))
if (length(need_names)) {
  dbn <- AnnotationDbi::select(GO.db, keys = need_names, keytype = "GOID", columns = "TERM")
  dbn <- dplyr::rename(dbn, go_id = GOID, go_name_db = TERM)
  go_map <- go_map %>%
    dplyr::left_join(dbn, by = "go_id") %>%
    dplyr::mutate(go_name = dplyr::coalesce(go_name, go_name_db)) %>%
    dplyr::select(-go_name_db)
}

## --- read shared DE genes, auto-detect ID column ---
stopifnot(file.exists(de_csv_path))
de_df <- readr::read_csv(de_csv_path, show_col_types = FALSE)

guess_id_col <- function(df){
  # prefer obvious columns
  cand <- intersect(names(df), c("gene_id","Gene ID","ID","Genes","NCU","NCU_ID","NCUid","Attributes","VALUE"))
  if (length(cand)) return(cand[1])
  # else pick column with most NCU-like entries
  score <- sapply(df, function(x) sum(grepl("^NCU\\d{5}$", as.character(x))))
  names(which.max(score))
}
de_col <- guess_id_col(de_df)
genes_shared <- unique(as.character(de_df[[de_col]]))

## --- universe: all genes that have any GO term (quick + conservative) ---
universe_genes <- unique(as.character(go_map$gene_id))

## --- TERM2GENE / TERM2NAME for Biological Process only ---
TERM2GENE <- go_map %>%
  dplyr::filter(tolower(namespace) %in% c("bp","biological process","biological_process","p")) %>%
  dplyr::select(go_id, gene_id) %>% dplyr::distinct()
TERM2NAME <- go_map %>%
  dplyr::select(go_id, go_name) %>% dplyr::distinct()

## --- enrichment (BH FDR; we’ll filter to FDR <= 0.05 for plotting) ---
enr_bp <- clusterProfiler::enricher(
  gene          = genes_shared,
  universe      = universe_genes,
  TERM2GENE     = TERM2GENE,
  TERM2NAME     = TERM2NAME,
  pAdjustMethod = "BH",
  pvalueCutoff  = 1.0,
  qvalueCutoff  = 1.0
)
if (is.null(enr_bp) || nrow(enr_bp@result) == 0)
  stop("No BP terms returned. Check IDs vs GO mapping.")

## --- compute fold-enrichment and tidy ---
res <- enr_bp@result %>%
  dplyr::mutate(
    GeneRatioNum = sapply(strsplit(GeneRatio, "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    BgRatioNum   = sapply(strsplit(BgRatio,   "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    FoldEnrich   = GeneRatioNum / BgRatioNum,
    FDR          = p.adjust(pvalue, "BH"),
    minusLog10FDR = -log10(FDR),
    Term = dplyr::coalesce(Description, ID)   # Description is the term name from TERM2NAME
  ) %>%
  dplyr::arrange(dplyr::desc(FoldEnrich))

readr::write_tsv(res, out_res)

## --- lollipop plot: names on y, bigger size spread, vivid colors; FDR-filtered ---
max_chars <- 37
shorten <- function(x, n = max_chars) ifelse(nchar(x) > n, paste0(substr(x, 1, n - 1), "…"), x)

plot_df <- res %>%
  dplyr::filter(FDR <= 0.05) %>%                # significant terms only
  dplyr::slice_head(n = top_n_terms) %>%
  dplyr::mutate(
    LabelShow = shorten(Term),                  # what you SEE (no GO IDs)
    LabelKey  = paste0(Term, "\u200B", dplyr::row_number()), # invisible uniqueness
    Label     = factor(LabelKey, levels = rev(LabelKey))
  )

lab_map <- setNames(plot_df$LabelShow, plot_df$LabelKey)

p <- ggplot(plot_df, aes(FoldEnrich, Label)) +
  geom_segment(aes(x = 0, xend = FoldEnrich, yend = Label), linewidth = 0.8) +
  geom_point(aes(size = Count, color = minusLog10FDR)) +
  scale_size(range = c(3, 10), name = "N. of Genes") +
  scale_color_viridis_c(option = "plasma", name = expression(-log[10]("FDR"))) +
  scale_y_discrete(labels = lab_map) +          # show shortened names only
  labs(x = "Fold Enrichment", y = NULL,
       title = expression("GO Biological Process Enrichment (cac-1suz-12 UP)")) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y  = element_text(size = 6),
    axis.text.x  = element_text(size = 9),
    legend.text  = element_text(size = 9),
    legend.title = element_text(size = 10),
    plot.title   = element_text(size = 12, hjust = 0.5),
    panel.grid.major.y = element_blank()
  )

ggplot2::ggsave(out_plot, p, width = 7, height = 5)
p


#########################################################################################

# GO BP enrichment + lollipop plot for Δcac-1Δsuz-12 DOWN genes
# Requires:
#   - Nc_GO_mapping.txt
#   - cac1_suz12__alpha0.05__DOWN_genes.csv

## --- paths / outputs ---
fungidb_go_path <- "Nc_GO_mapping.txt"
de_csv_path     <- "../CAF-1_RNA-seq_Analysis/csv_files/cac1_suz12__alpha0.05__DOWN_genes.csv"
out_tag  <- "cac1_suz12_DOWN_GO"
out_res  <- paste0(out_tag, "_BP_enrichment.tsv")
out_plot <- paste0(out_tag, "_BP_lollipop.pdf")
top_n_terms <- 20

## --- read GO mapping (FungiDB tab with separate ID/name cols) ---
stopifnot(file.exists(fungidb_go_path))
go_raw <- readr::read_tsv(fungidb_go_path, show_col_types = FALSE)

gene_col <- intersect(names(go_raw), c("Gene ID","gene_id","Genes","ID"))[1]
if (is.na(gene_col)) stop("Could not find a 'Gene ID' column in Nc_GO_mapping.txt")

pairs <- list(
  list(id="Curated GO Process IDs",   nm="Curated GO Processes",   ns="BP"),
  list(id="Computed GO Process IDs",  nm="Computed GO Processes",   ns="BP"),
  list(id="Curated GO Component IDs", nm="Curated GO Components",   ns="CC"),
  list(id="Computed GO Component IDs",nm="Computed GO Components",  ns="CC"),
  list(id="Curated GO Function IDs",  nm="Curated GO Functions",    ns="MF"),
  list(id="Computed GO Function IDs", nm="Computed GO Functions",   ns="MF")
)

make_long <- function(idcol, nmcol, ns){
  if (!(idcol %in% names(go_raw))) return(NULL)
  tmp <- go_raw %>%
    dplyr::transmute(
      gene_id = .data[[gene_col]],
      go_id   = as.character(.data[[idcol]]),
      go_name = if (nmcol %in% names(go_raw)) as.character(.data[[nmcol]]) else NA_character_
    ) %>%
    dplyr::filter(!is.na(go_id), nzchar(go_id)) %>%
    tidyr::separate_rows(go_id,  sep = "\\s*;\\s*") %>%
    tidyr::separate_rows(go_name, sep = "\\s*;\\s*") %>%
    dplyr::mutate(namespace = ns) %>%
    dplyr::distinct()
  tmp
}

go_map <- do.call(
  dplyr::bind_rows,
  lapply(pairs, function(p) make_long(p$id, p$nm, p$ns))
) %>%
  dplyr::filter(grepl("^GO:", go_id)) %>%
  dplyr::distinct()

# fill missing names from GO.db
need_names <- setdiff(unique(go_map$go_id), unique(go_map$go_id[!is.na(go_map$go_name) & nzchar(go_map$go_name)]))
if (length(need_names)) {
  dbn <- AnnotationDbi::select(GO.db, keys = need_names, keytype = "GOID", columns = "TERM")
  dbn <- dplyr::rename(dbn, go_id = GOID, go_name_db = TERM)
  go_map <- go_map %>%
    dplyr::left_join(dbn, by = "go_id") %>%
    dplyr::mutate(go_name = dplyr::coalesce(go_name, go_name_db)) %>%
    dplyr::select(-go_name_db)
}

## --- read shared DE genes, auto-detect ID column ---
stopifnot(file.exists(de_csv_path))
de_df <- readr::read_csv(de_csv_path, show_col_types = FALSE)

guess_id_col <- function(df){
  # prefer obvious columns
  cand <- intersect(names(df), c("gene_id","Gene ID","ID","Genes","NCU","NCU_ID","NCUid","Attributes","VALUE"))
  if (length(cand)) return(cand[1])
  # else pick column with most NCU-like entries
  score <- sapply(df, function(x) sum(grepl("^NCU\\d{5}$", as.character(x))))
  names(which.max(score))
}
de_col <- guess_id_col(de_df)
genes_shared <- unique(as.character(de_df[[de_col]]))

## --- universe: all genes that have any GO term (quick + conservative) ---
universe_genes <- unique(as.character(go_map$gene_id))

## --- TERM2GENE / TERM2NAME for Biological Process only ---
TERM2GENE <- go_map %>%
  dplyr::filter(tolower(namespace) %in% c("bp","biological process","biological_process","p")) %>%
  dplyr::select(go_id, gene_id) %>% dplyr::distinct()
TERM2NAME <- go_map %>%
  dplyr::select(go_id, go_name) %>% dplyr::distinct()

## --- enrichment (BH FDR; we’ll filter to FDR <= 0.05 for plotting) ---
enr_bp <- clusterProfiler::enricher(
  gene          = genes_shared,
  universe      = universe_genes,
  TERM2GENE     = TERM2GENE,
  TERM2NAME     = TERM2NAME,
  pAdjustMethod = "BH",
  pvalueCutoff  = 1.0,
  qvalueCutoff  = 1.0
)
if (is.null(enr_bp) || nrow(enr_bp@result) == 0)
  stop("No BP terms returned. Check IDs vs GO mapping.")

## --- compute fold-enrichment and tidy ---
res <- enr_bp@result %>%
  dplyr::mutate(
    GeneRatioNum   = sapply(strsplit(GeneRatio, "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    BgRatioNum     = sapply(strsplit(BgRatio,   "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    FoldEnrich     = GeneRatioNum / BgRatioNum,
    FDR            = p.adjust(pvalue, "BH"),
    minusLog10FDR  = -log10(FDR),
    Term           = dplyr::coalesce(Description, ID)   # use GO term name (no GO ID)
  ) %>%
  dplyr::arrange(dplyr::desc(FoldEnrich))

readr::write_tsv(res, out_res)

## ---- plotting: show ONLY names, shorten long labels, keep uniqueness invisibly ----
max_chars <- 37
shorten <- function(x, n = max_chars) ifelse(nchar(x) > n, paste0(substr(x, 1, n - 1), "…"), x)

plot_df <- res %>%
  dplyr::filter(FDR <= 0.05) %>%                # significant terms only
  dplyr::slice_head(n = top_n_terms) %>%
  dplyr::mutate(
    LabelShow = shorten(Term),                  # what you SEE (no GO IDs)
    LabelKey  = paste0(Term, "\u200B", dplyr::row_number()), # invisible uniqueness
    Label     = factor(LabelKey, levels = rev(LabelKey))
  )

lab_map <- setNames(plot_df$LabelShow, plot_df$LabelKey)

p <- ggplot(plot_df, aes(FoldEnrich, Label)) +
  geom_segment(aes(x = 0, xend = FoldEnrich, yend = Label), linewidth = 0.8) +
  geom_point(aes(size = Count, color = minusLog10FDR)) +
  scale_size(range = c(3, 10), name = "N. of Genes") +
  scale_color_viridis_c(option = "plasma", name = expression(-log[10]("FDR"))) +
  scale_y_discrete(labels = lab_map) +          # show shortened names only
  labs(x = "Fold Enrichment", y = NULL,
       title = expression("GO Biological Process Enrichment (cac-1suz-12 DOWN)")) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y  = element_text(size = 6),
    axis.text.x  = element_text(size = 9),
    legend.text  = element_text(size = 9),
    legend.title = element_text(size = 10),
    plot.title   = element_text(size = 12, hjust = 0.5),
    panel.grid.major.y = element_blank()
  )

ggplot2::ggsave(out_plot, p, width = 7, height = 5)
p

########################################################################################
###GO Analysis on H3K27me3-marked genes
all_genes <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac1__alpha0.05__ALL_results_LFC.csv")
K27genes  <- subset(all_genes, all_genes$NCU %in% rownames(AVERAGE_Prc2targetTPM))

## --- paths / outputs ---
fungidb_go_path <- "Nc_GO_mapping.txt"
de_input <- K27genes
out_tag <- "PRC2_targets_GO"
out_res <- paste0(out_tag, "_BP_enrichment.tsv")
out_plot <- paste0(out_tag, "_BP_lollipop.pdf")
top_n_terms <- 20

## --- read GO mapping (FungiDB tab with separate ID/name cols) ---
stopifnot(file.exists(fungidb_go_path))
go_raw <- readr::read_tsv(fungidb_go_path, show_col_types = FALSE)

gene_col <- intersect(names(go_raw), c("Gene ID","gene_id","Genes","ID"))[1]
if (is.na(gene_col)) stop("Could not find a 'Gene ID' column in Nc_GO_mapping.txt")

pairs <- list(
  list(id="Curated GO Process IDs",   nm="Curated GO Processes",   ns="BP"),
  list(id="Computed GO Process IDs",  nm="Computed GO Processes",   ns="BP"),
  list(id="Curated GO Component IDs", nm="Curated GO Components",   ns="CC"),
  list(id="Computed GO Component IDs",nm="Computed GO Components",  ns="CC"),
  list(id="Curated GO Function IDs",  nm="Curated GO Functions",    ns="MF"),
  list(id="Computed GO Function IDs", nm="Computed GO Functions",   ns="MF")
)

make_long <- function(idcol, nmcol, ns){
  if (!(idcol %in% names(go_raw))) return(NULL)
  tmp <- go_raw %>%
    dplyr::transmute(
      gene_id = .data[[gene_col]],
      go_id   = as.character(.data[[idcol]]),
      go_name = if (nmcol %in% names(go_raw)) as.character(.data[[nmcol]]) else NA_character_
    ) %>%
    dplyr::filter(!is.na(go_id), nzchar(go_id)) %>%
    tidyr::separate_rows(go_id,  sep = "\\s*;\\s*") %>%
    tidyr::separate_rows(go_name, sep = "\\s*;\\s*") %>%
    dplyr::mutate(namespace = ns) %>%
    dplyr::distinct()
  tmp
}

go_map <- do.call(
  dplyr::bind_rows,
  lapply(pairs, function(p) make_long(p$id, p$nm, p$ns))
) %>%
  dplyr::filter(grepl("^GO:", go_id)) %>%
  dplyr::distinct()

# fill missing names from GO.db
need_names <- setdiff(unique(go_map$go_id), unique(go_map$go_id[!is.na(go_map$go_name) & nzchar(go_map$go_name)]))
if (length(need_names)) {
  dbn <- AnnotationDbi::select(GO.db, keys = need_names, keytype = "GOID", columns = "TERM")
  dbn <- dplyr::rename(dbn, go_id = GOID, go_name_db = TERM)
  go_map <- go_map %>%
    dplyr::left_join(dbn, by = "go_id") %>%
    dplyr::mutate(go_name = dplyr::coalesce(go_name, go_name_db)) %>%
    dplyr::select(-go_name_db)
}

## --- read shared DE genes, auto-detect ID column ---
get_de_df <- function(x) {
  if (is.data.frame(x) || inherits(x, "tbl_df")) {
    return(x)
  }
  if (is.character(x) && length(x) == 1L) {
    if (!file.exists(x)) {
      stop("File does not exist: ", x)
    }
    return(readr::read_csv(x, show_col_types = FALSE))
  }
  stop("de_input must be either a data.frame or a single file path string.")
}

de_df <- get_de_df(de_input)

guess_id_col <- function(df){
  # prefer obvious columns
  cand <- intersect(names(df), c("gene_id","Gene ID","ID","Genes","NCU","NCU_ID","NCUid","Attributes","VALUE"))
  if (length(cand)) return(cand[1])
  # else pick column with most NCU-like entries
  score <- sapply(df, function(x) sum(grepl("^NCU\\d{5}$", as.character(x))))
  names(which.max(score))
}
de_col <- guess_id_col(de_df)
genes_shared <- unique(as.character(de_df[[de_col]]))

## --- universe: all genes that have any GO term (quick + conservative) ---
universe_genes <- unique(as.character(go_map$gene_id))

## --- TERM2GENE / TERM2NAME for Biological Process only ---
TERM2GENE <- go_map %>%
  dplyr::filter(tolower(namespace) %in% c("bp","biological process","biological_process","p")) %>%
  dplyr::select(go_id, gene_id) %>% dplyr::distinct()
TERM2NAME <- go_map %>%
  dplyr::select(go_id, go_name) %>% dplyr::distinct()

## --- enrichment (BH FDR; we’ll filter to FDR <= 0.05 for plotting) ---
enr_bp <- clusterProfiler::enricher(
  gene          = genes_shared,
  universe      = universe_genes,
  TERM2GENE     = TERM2GENE,
  TERM2NAME     = TERM2NAME,
  pAdjustMethod = "BH",
  pvalueCutoff  = 1.0,
  qvalueCutoff  = 1.0
)
if (is.null(enr_bp) || nrow(enr_bp@result) == 0)
  stop("No BP terms returned. Check IDs vs GO mapping.")

## --- compute fold-enrichment and tidy ---
res <- enr_bp@result %>%
  dplyr::mutate(
    GeneRatioNum = sapply(strsplit(GeneRatio, "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    BgRatioNum   = sapply(strsplit(BgRatio,   "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    FoldEnrich   = GeneRatioNum / BgRatioNum,
    FDR          = p.adjust(pvalue, "BH"),
    minusLog10FDR = -log10(FDR),
    Term = dplyr::coalesce(Description, ID)
  ) %>%
  dplyr::arrange(dplyr::desc(FoldEnrich))

readr::write_tsv(res, out_res)

## --- lollipop plot: names on y, bigger size spread, vivid colors; FDR-filtered ---
max_chars <- 37
shorten <- function(x, n = max_chars) ifelse(nchar(x) > n, paste0(substr(x, 1, n - 1), "…"), x)

plot_df <- res %>%
  dplyr::filter(FDR <= 0.05) %>%                # significant terms only
  dplyr::slice_head(n = top_n_terms) %>%
  dplyr::mutate(
    LabelShow = shorten(Term),                  # what you SEE (no GO IDs)
    LabelKey  = paste0(Term, "\u200B", dplyr::row_number()), # invisible uniqueness
    Label     = factor(LabelKey, levels = rev(LabelKey))
  )

lab_map <- setNames(plot_df$LabelShow, plot_df$LabelKey)

p <- ggplot(plot_df, aes(FoldEnrich, Label)) +
  geom_segment(aes(x = 0, xend = FoldEnrich, yend = Label), linewidth = 0.8) +
  geom_point(aes(size = Count, color = minusLog10FDR)) +
  scale_size(range = c(3, 10), name = "N. of Genes") +
  scale_color_viridis_c(option = "plasma", name = expression(-log[10]("FDR"))) +
  scale_y_discrete(labels = lab_map) +
  labs(x = "Fold Enrichment", y = NULL,
       title = expression("GO Biological Process Enrichment H3K27me3-marked genes")) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y  = element_text(size = 6),
    axis.text.x  = element_text(size = 9),
    legend.text  = element_text(size = 9),
    legend.title = element_text(size = 10),
    plot.title   = element_text(size = 12, hjust = 0.5),
    panel.grid.major.y = element_blank()
  )

ggplot2::ggsave(out_plot, p, width = 7, height = 5)
p

########################################################################################
###GO Analysis on genes upregulated in set-7
FP_DE_genes <- read.table(file = "../text_files/FP_DE_genes.txt", header=TRUE, stringsAsFactors=FALSE, row.names=NULL, check.names=FALSE, sep="\t")

## --- paths / outputs ---
fungidb_go_path <- "Nc_GO_mapping.txt"
de_input <- FP_DE_genes
out_tag <- "FP_DE_GO"
out_res <- paste0(out_tag, "_BP_enrichment.tsv")
out_plot <- paste0(out_tag, "_BP_lollipop.pdf")
top_n_terms <- 20

## --- read GO mapping (FungiDB tab with separate ID/name cols) ---
stopifnot(file.exists(fungidb_go_path))
go_raw <- readr::read_tsv(fungidb_go_path, show_col_types = FALSE)

gene_col <- intersect(names(go_raw), c("Gene ID","gene_id","Genes","ID"))[1]
if (is.na(gene_col)) stop("Could not find a 'Gene ID' column in Nc_GO_mapping.txt")

pairs <- list(
  list(id="Curated GO Process IDs",   nm="Curated GO Processes",   ns="BP"),
  list(id="Computed GO Process IDs",  nm="Computed GO Processes",   ns="BP"),
  list(id="Curated GO Component IDs", nm="Curated GO Components",   ns="CC"),
  list(id="Computed GO Component IDs",nm="Computed GO Components",  ns="CC"),
  list(id="Curated GO Function IDs",  nm="Curated GO Functions",    ns="MF"),
  list(id="Computed GO Function IDs", nm="Computed GO Functions",   ns="MF")
)

make_long <- function(idcol, nmcol, ns){
  if (!(idcol %in% names(go_raw))) return(NULL)
  tmp <- go_raw %>%
    dplyr::transmute(
      gene_id = .data[[gene_col]],
      go_id   = as.character(.data[[idcol]]),
      go_name = if (nmcol %in% names(go_raw)) as.character(.data[[nmcol]]) else NA_character_
    ) %>%
    dplyr::filter(!is.na(go_id), nzchar(go_id)) %>%
    tidyr::separate_rows(go_id,  sep = "\\s*;\\s*") %>%
    tidyr::separate_rows(go_name, sep = "\\s*;\\s*") %>%
    dplyr::mutate(namespace = ns) %>%
    dplyr::distinct()
  tmp
}

go_map <- do.call(
  dplyr::bind_rows,
  lapply(pairs, function(p) make_long(p$id, p$nm, p$ns))
) %>%
  dplyr::filter(grepl("^GO:", go_id)) %>%
  dplyr::distinct()

# fill missing names from GO.db
need_names <- setdiff(unique(go_map$go_id), unique(go_map$go_id[!is.na(go_map$go_name) & nzchar(go_map$go_name)]))
if (length(need_names)) {
  dbn <- AnnotationDbi::select(GO.db, keys = need_names, keytype = "GOID", columns = "TERM")
  dbn <- dplyr::rename(dbn, go_id = GOID, go_name_db = TERM)
  go_map <- go_map %>%
    dplyr::left_join(dbn, by = "go_id") %>%
    dplyr::mutate(go_name = dplyr::coalesce(go_name, go_name_db)) %>%
    dplyr::select(-go_name_db)
}

## --- read shared DE genes, auto-detect ID column ---
get_de_df <- function(x) {
  if (is.data.frame(x) || inherits(x, "tbl_df")) {
    return(x)
  }
  if (is.character(x) && length(x) == 1L) {
    if (!file.exists(x)) {
      stop("File does not exist: ", x)
    }
    return(readr::read_csv(x, show_col_types = FALSE))
  }
  stop("de_input must be either a data.frame or a single file path string.")
}

de_df <- get_de_df(de_input)

guess_id_col <- function(df){
  # prefer obvious columns
  cand <- intersect(names(df), c("gene_id","Gene ID","ID","Genes","NCU","NCU_ID","NCUid","Attributes","VALUE"))
  if (length(cand)) return(cand[1])
  # else pick column with most NCU-like entries
  score <- sapply(df, function(x) sum(grepl("^NCU\\d{5}$", as.character(x))))
  names(which.max(score))
}
de_col <- guess_id_col(de_df)
genes_shared <- unique(as.character(de_df[[de_col]]))

## --- universe: all genes that have any GO term (quick + conservative) ---
universe_genes <- unique(as.character(go_map$gene_id))

## --- TERM2GENE / TERM2NAME for Biological Process only ---
TERM2GENE <- go_map %>%
  dplyr::filter(tolower(namespace) %in% c("bp","biological process","biological_process","p")) %>%
  dplyr::select(go_id, gene_id) %>% dplyr::distinct()
TERM2NAME <- go_map %>%
  dplyr::select(go_id, go_name) %>% dplyr::distinct()

## --- enrichment (BH FDR; we’ll filter to FDR <= 0.05 for plotting) ---
enr_bp <- clusterProfiler::enricher(
  gene          = genes_shared,
  universe      = universe_genes,
  TERM2GENE     = TERM2GENE,
  TERM2NAME     = TERM2NAME,
  pAdjustMethod = "BH",
  pvalueCutoff  = 1.0,
  qvalueCutoff  = 1.0
)
if (is.null(enr_bp) || nrow(enr_bp@result) == 0)
  stop("No BP terms returned. Check IDs vs GO mapping.")

## --- compute fold-enrichment and tidy ---
res <- enr_bp@result %>%
  dplyr::mutate(
    GeneRatioNum = sapply(strsplit(GeneRatio, "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    BgRatioNum   = sapply(strsplit(BgRatio,   "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    FoldEnrich   = GeneRatioNum / BgRatioNum,
    FDR          = p.adjust(pvalue, "BH"),
    minusLog10FDR = -log10(FDR),
    Term = dplyr::coalesce(Description, ID)
  ) %>%
  dplyr::arrange(dplyr::desc(FoldEnrich))

readr::write_tsv(res, out_res)

## --- lollipop plot: names on y, bigger size spread, vivid colors; FDR-filtered ---
max_chars <- 37
shorten <- function(x, n = max_chars) ifelse(nchar(x) > n, paste0(substr(x, 1, n - 1), "…"), x)

plot_df <- res %>%
  dplyr::filter(FDR <= 0.05) %>%                # significant terms only
  dplyr::slice_head(n = top_n_terms) %>%
  dplyr::mutate(
    LabelShow = shorten(Term),                  # what you SEE (no GO IDs)
    LabelKey  = paste0(Term, "\u200B", dplyr::row_number()), # invisible uniqueness
    Label     = factor(LabelKey, levels = rev(LabelKey))
  )

lab_map <- setNames(plot_df$LabelShow, plot_df$LabelKey)

p <- ggplot(plot_df, aes(FoldEnrich, Label)) +
  geom_segment(aes(x = 0, xend = FoldEnrich, yend = Label), linewidth = 0.8) +
  geom_point(aes(size = Count, color = minusLog10FDR)) +
  scale_size(range = c(3, 10), name = "N. of Genes") +
  scale_color_viridis_c(option = "plasma", name = expression(-log[10]("FDR"))) +
  scale_y_discrete(labels = lab_map) +
  labs(x = "Fold Enrichment", y = NULL,
       title = expression("GO Biological Process Enrichment DE in False Perithecia")) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y  = element_text(size = 6),
    axis.text.x  = element_text(size = 9),
    legend.text  = element_text(size = 9),
    legend.title = element_text(size = 10),
    plot.title   = element_text(size = 12, hjust = 0.5),
    panel.grid.major.y = element_blank()
  )

ggplot2::ggsave(out_plot, p, width = 7, height = 5)
p

########################################################################################
###GO Analysis on genes upregulated in set-7
set7_DE_genes <- read.table(file = "../text_files/set7_DE_genes.txt", header=TRUE, stringsAsFactors=FALSE, row.names=NULL, check.names=FALSE, sep="\t")

## --- paths / outputs ---
fungidb_go_path <- "Nc_GO_mapping.txt"
de_input <- set7_DE_genes
out_tag <- "set-7_DE_GO"
out_res <- paste0(out_tag, "_BP_enrichment.tsv")
out_plot <- paste0(out_tag, "_BP_lollipop.pdf")
top_n_terms <- 20

## --- read GO mapping (FungiDB tab with separate ID/name cols) ---
stopifnot(file.exists(fungidb_go_path))
go_raw <- readr::read_tsv(fungidb_go_path, show_col_types = FALSE)

gene_col <- intersect(names(go_raw), c("Gene ID","gene_id","Genes","ID"))[1]
if (is.na(gene_col)) stop("Could not find a 'Gene ID' column in Nc_GO_mapping.txt")

pairs <- list(
  list(id="Curated GO Process IDs",   nm="Curated GO Processes",   ns="BP"),
  list(id="Computed GO Process IDs",  nm="Computed GO Processes",   ns="BP"),
  list(id="Curated GO Component IDs", nm="Curated GO Components",   ns="CC"),
  list(id="Computed GO Component IDs",nm="Computed GO Components",  ns="CC"),
  list(id="Curated GO Function IDs",  nm="Curated GO Functions",    ns="MF"),
  list(id="Computed GO Function IDs", nm="Computed GO Functions",   ns="MF")
)

make_long <- function(idcol, nmcol, ns){
  if (!(idcol %in% names(go_raw))) return(NULL)
  tmp <- go_raw %>%
    dplyr::transmute(
      gene_id = .data[[gene_col]],
      go_id   = as.character(.data[[idcol]]),
      go_name = if (nmcol %in% names(go_raw)) as.character(.data[[nmcol]]) else NA_character_
    ) %>%
    dplyr::filter(!is.na(go_id), nzchar(go_id)) %>%
    tidyr::separate_rows(go_id,  sep = "\\s*;\\s*") %>%
    tidyr::separate_rows(go_name, sep = "\\s*;\\s*") %>%
    dplyr::mutate(namespace = ns) %>%
    dplyr::distinct()
  tmp
}

go_map <- do.call(
  dplyr::bind_rows,
  lapply(pairs, function(p) make_long(p$id, p$nm, p$ns))
) %>%
  dplyr::filter(grepl("^GO:", go_id)) %>%
  dplyr::distinct()

# fill missing names from GO.db
need_names <- setdiff(unique(go_map$go_id), unique(go_map$go_id[!is.na(go_map$go_name) & nzchar(go_map$go_name)]))
if (length(need_names)) {
  dbn <- AnnotationDbi::select(GO.db, keys = need_names, keytype = "GOID", columns = "TERM")
  dbn <- dplyr::rename(dbn, go_id = GOID, go_name_db = TERM)
  go_map <- go_map %>%
    dplyr::left_join(dbn, by = "go_id") %>%
    dplyr::mutate(go_name = dplyr::coalesce(go_name, go_name_db)) %>%
    dplyr::select(-go_name_db)
}

## --- read shared DE genes, auto-detect ID column ---
get_de_df <- function(x) {
  if (is.data.frame(x) || inherits(x, "tbl_df")) {
    return(x)
  }
  if (is.character(x) && length(x) == 1L) {
    if (!file.exists(x)) {
      stop("File does not exist: ", x)
    }
    return(readr::read_csv(x, show_col_types = FALSE))
  }
  stop("de_input must be either a data.frame or a single file path string.")
}

de_df <- get_de_df(de_input)

guess_id_col <- function(df){
  # prefer obvious columns
  cand <- intersect(names(df), c("gene_id","Gene ID","ID","Genes","NCU","NCU_ID","NCUid","Attributes","VALUE"))
  if (length(cand)) return(cand[1])
  # else pick column with most NCU-like entries
  score <- sapply(df, function(x) sum(grepl("^NCU\\d{5}$", as.character(x))))
  names(which.max(score))
}
de_col <- guess_id_col(de_df)
genes_shared <- unique(as.character(de_df[[de_col]]))

## --- universe: all genes that have any GO term (quick + conservative) ---
universe_genes <- unique(as.character(go_map$gene_id))

## --- TERM2GENE / TERM2NAME for Biological Process only ---
TERM2GENE <- go_map %>%
  dplyr::filter(tolower(namespace) %in% c("bp","biological process","biological_process","p")) %>%
  dplyr::select(go_id, gene_id) %>% dplyr::distinct()
TERM2NAME <- go_map %>%
  dplyr::select(go_id, go_name) %>% dplyr::distinct()

## --- enrichment (BH FDR; we’ll filter to FDR <= 0.05 for plotting) ---
enr_bp <- clusterProfiler::enricher(
  gene          = genes_shared,
  universe      = universe_genes,
  TERM2GENE     = TERM2GENE,
  TERM2NAME     = TERM2NAME,
  pAdjustMethod = "BH",
  pvalueCutoff  = 1.0,
  qvalueCutoff  = 1.0
)
if (is.null(enr_bp) || nrow(enr_bp@result) == 0)
  stop("No BP terms returned. Check IDs vs GO mapping.")

## --- compute fold-enrichment and tidy ---
res <- enr_bp@result %>%
  dplyr::mutate(
    GeneRatioNum = sapply(strsplit(GeneRatio, "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    BgRatioNum   = sapply(strsplit(BgRatio,   "/"), function(x) as.numeric(x[1]) / as.numeric(x[2])),
    FoldEnrich   = GeneRatioNum / BgRatioNum,
    FDR          = p.adjust(pvalue, "BH"),
    minusLog10FDR = -log10(FDR),
    Term = dplyr::coalesce(Description, ID)
  ) %>%
  dplyr::arrange(dplyr::desc(FoldEnrich))

readr::write_tsv(res, out_res)

## --- lollipop plot: names on y, bigger size spread, vivid colors; FDR-filtered ---
max_chars <- 37
shorten <- function(x, n = max_chars) ifelse(nchar(x) > n, paste0(substr(x, 1, n - 1), "…"), x)

plot_df <- res %>%
  dplyr::filter(FDR <= 0.05) %>%                # significant terms only
  dplyr::slice_head(n = top_n_terms) %>%
  dplyr::mutate(
    LabelShow = shorten(Term),                  # what you SEE (no GO IDs)
    LabelKey  = paste0(Term, "\u200B", dplyr::row_number()), # invisible uniqueness
    Label     = factor(LabelKey, levels = rev(LabelKey))
  )

lab_map <- setNames(plot_df$LabelShow, plot_df$LabelKey)

p <- ggplot(plot_df, aes(FoldEnrich, Label)) +
  geom_segment(aes(x = 0, xend = FoldEnrich, yend = Label), linewidth = 0.8) +
  geom_point(aes(size = Count, color = minusLog10FDR)) +
  scale_size(range = c(3, 10), name = "N. of Genes") +
  scale_color_viridis_c(option = "plasma", name = expression(-log[10]("FDR"))) +
  scale_y_discrete(labels = lab_map) +
  labs(x = "Fold Enrichment", y = NULL,
       title = expression("GO Biological Process Enrichment Upregulated in set-7")) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y  = element_text(size = 6),
    axis.text.x  = element_text(size = 9),
    legend.text  = element_text(size = 9),
    legend.title = element_text(size = 10),
    plot.title   = element_text(size = 12, hjust = 0.5),
    panel.grid.major.y = element_blank()
  )

ggplot2::ggsave(out_plot, p, width = 7, height = 5)
p

```

#Next, make Volcano Plots displaying all DE genes
```{r, Volcano Plots}
#Make New variables for plot
##read csvs with log2FC data
cac1_vol <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac1__alpha0.05__ALL_results_LFC.csv")
cac2_vol <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac2__alpha0.05__ALL_results_LFC.csv")
cac3_vol <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac3__alpha0.05__ALL_results_LFC.csv")
set7_vol <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/set7__alpha0.05__ALL_results_LFC.csv")
cac1_suz12_vol <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac1_SUZ12__alpha0.05__ALL_results_LFC.csv")

#I need to get rid of a few genes with a p-value of 0, as well as the knocked-out gene itself
cac1_vol <- cac1_vol[cac1_vol$padj != 0, ]
cac2_vol <- cac2_vol[cac2_vol$padj != 0, ]
cac3_vol <- cac3_vol[cac3_vol$padj != 0, ]
set7_vol <- set7_vol[set7_vol$padj != 0, ]
cac1_suz12_vol <- cac1_suz12_vol[cac1_vol$padj != 0, ]
cac1_vol <- cac1_vol[-which.min(cac1_vol$log2FoldChange), ]
cac2_vol <- cac2_vol[-which.min(cac2_vol$log2FoldChange), ]
cac3_vol <- cac3_vol[-which.min(cac3_vol$log2FoldChange), ]
set7_vol <- set7_vol[-which.min(set7_vol$log2FoldChange), ]
cac1_suz12_vol <- cac1_suz12_vol[-which.min(cac1_suz12_vol$log2FoldChange), ]
cac1_suz12_vol <- subset(cac1_suz12_vol, NCU != "NCU05460")

##cac-1
#Assign categories with NS prioritized
cac1_vol$category <- case_when(
  cac1_vol$log2FoldChange > -1 & cac1_vol$log2FoldChange < 1 ~ "NS",
  cac1_vol$NCU %in% Prc2targets$V11 ~ "PRC2",
  cac1_vol$log2FoldChange > 1 & cac1_vol$padj < 0.05 ~ "Upregulated",
  cac1_vol$log2FoldChange < -1 & cac1_vol$padj < 0.05 ~ "Downregulated",
  TRUE ~ "NS"
)

#Reorder so PRC2 is plotted last (on top)
cac1_vol <- cac1_vol[order(cac1_vol$category == "PRC2"), ]

#Plot
library(ggplot2)
library(scales)  # breaks_width(), pseudo_log_trans()

cac1_volcano_plot <- ggplot(cac1_vol, aes(
  x = log2FoldChange,
  y = -log10(padj),
  color = category
)) +
  geom_point(size = 3, alpha = 0.85) +
  scale_color_manual(
    values = c(
      "Upregulated" = "red",
      "Downregulated" = "blue",
      "PRC2" = "orange2",
      "NS" = "grey"
    ),
    breaks = c("Upregulated", "Downregulated", "PRC2"), # only show these in legend
    labels = c("Upregulated", "Downregulated", "PRC2")
  ) +
  labs(
    x = expression(paste("log"[2], "FoldChange(cac-1/WT)")),
    y = expression(paste("-log"[10], "(p-value)")),
    title = "cac-1 vs WT"
  ) + 
  scale_x_continuous(breaks = breaks_width(3), minor_breaks = breaks_width(1)) + 
  scale_y_continuous(trans = pseudo_log_trans(base = 10, sigma = 2),
                     breaks = c(0,2,5,10,20,50,100,200,300),
                     limits = c(0, NA)) + 
  coord_cartesian(ylim = c(0, 300)) +
  geom_vline(xintercept = c(-1, 1), color = "grey5", linetype = "dashed", alpha = 0.35) +
  geom_hline(yintercept = -log10(0.05), color = "grey5", linetype = "dashed", alpha = 0.35) +
  theme_classic(base_size = 20) +
  theme(
    axis.title.y = element_text(face = "bold", margin = margin(0, 20, 0, 0), size = rel(1.1), color = 'black'),
    axis.title.x = element_text(face = "bold", margin = margin(20, 0, 0, 0), size = rel(1.1), color = 'black',),
    axis.text = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5, color = "black"),
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.title = element_blank() 
  ) +
  guides(color = guide_legend(override.aes = list(size = 5)))  # bigger legend dots

#Print the plot
print(cac1_volcano_plot)

ggsave("./cac1_volcano_plot_FINAL.pdf", plot=cac1_volcano_plot, width = 12, height = 9, unit="in",  dpi=400)

##cac-2
#Assign categories with NS prioritized
cac2_vol$category <- case_when(
  cac2_vol$log2FoldChange > -1 & cac2_vol$log2FoldChange < 1 ~ "NS",
  cac2_vol$NCU %in% Prc2targets$V11 ~ "PRC2",
  cac2_vol$log2FoldChange > 1 & cac2_vol$padj < 0.05 ~ "Upregulated",
  cac2_vol$log2FoldChange < -1 & cac2_vol$padj < 0.05 ~ "Downregulated",
  TRUE ~ "NS"
)

#Reorder so PRC2 is plotted last (on top)
cac2_vol <- cac2_vol[order(cac2_vol$category == "PRC2"), ]

#Plot
cac2_volcano_plot <- ggplot(cac2_vol, aes(
  x = log2FoldChange,
  y = -log10(padj),
  color = category
)) +
  geom_point(size = 3, alpha = 0.85) +
  scale_color_manual(
    values = c(
      "Upregulated" = "red",
      "Downregulated" = "blue",
      "PRC2" = "orange2",
      "NS" = "grey"
    ),
    breaks = c("Upregulated", "Downregulated", "PRC2"), # only show these in legend
    labels = c("Upregulated", "Downregulated", "PRC2")
  ) +
  labs(
    x = expression(paste("log"[2], "FoldChange(cac-2/WT)")),
    y = expression(paste("-log"[10], "(p-value)")),
    title = "cac-2 vs WT"
  ) + 
  scale_x_continuous(breaks = breaks_width(3), minor_breaks = breaks_width(1)) + 
  scale_y_continuous(trans = pseudo_log_trans(base = 10, sigma = 2),
                     breaks = c(0,2,5,10,20,50,100,200,300),
                     limits = c(0, NA)) + 
  coord_cartesian(ylim = c(0, 300)) +
  geom_vline(xintercept = c(-1, 1), color = "grey5", linetype = "dashed", alpha = 0.35) +
  geom_hline(yintercept = -log10(0.05), color = "grey5", linetype = "dashed", alpha = 0.35) +
  theme_classic(base_size = 20) +
  theme(
    axis.title.y = element_text(face = "bold", margin = margin(0, 20, 0, 0), size = rel(1.1), color = 'black'),
    axis.title.x = element_text(face = "bold", margin = margin(20, 0, 0, 0), size = rel(1.1), color = 'black',),
    axis.text = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5, color = "black"),
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.title = element_blank() 
  ) +
  guides(color = guide_legend(override.aes = list(size = 5)))  # bigger legend dots

#Print the plot
print(cac2_volcano_plot)

ggsave("./cac2_volcano_plot_FINAL.pdf", plot=cac2_volcano_plot, width = 12, height = 9, unit="in",  dpi=400)

##cac-3
#Assign categories with NS prioritized
cac3_vol$category <- case_when(
  cac3_vol$log2FoldChange > -1 & cac3_vol$log2FoldChange < 1 ~ "NS",
  cac3_vol$NCU %in% Prc2targets$V11 ~ "PRC2",
  cac3_vol$log2FoldChange > 1 & cac3_vol$padj < 0.05 ~ "Upregulated",
  cac3_vol$log2FoldChange < -1 & cac3_vol$padj < 0.05 ~ "Downregulated",
  TRUE ~ "NS"
)

#Reorder so PRC2 is plotted last (on top)
cac3_vol <- cac3_vol[order(cac3_vol$category == "PRC2"), ]

#Plot
cac3_volcano_plot <- ggplot(cac3_vol, aes(
  x = log2FoldChange,
  y = -log10(padj),
  color = category
)) +
  geom_point(size = 3, alpha = 0.85) +
  scale_color_manual(
    values = c(
      "Upregulated" = "red",
      "Downregulated" = "blue",
      "PRC2" = "orange2",
      "NS" = "grey"
    ),
    breaks = c("Upregulated", "Downregulated", "PRC2"), # only show these in legend
    labels = c("Upregulated", "Downregulated", "PRC2")
  ) +
  labs(
    x = expression(paste("log"[2], "FoldChange(cac-3/WT)")),
    y = expression(paste("-log"[10], "(p-value)")),
    title = "cac-3 vs WT"
  ) + 
  scale_x_continuous(breaks = breaks_width(3), minor_breaks = breaks_width(1)) + 
  scale_y_continuous(trans = pseudo_log_trans(base = 10, sigma = 2),
                     breaks = c(0,2,5,10,20,50,100,200,300),
                     limits = c(0, NA)) + 
  coord_cartesian(ylim = c(0, 300)) +
  geom_vline(xintercept = c(-1, 1), color = "grey5", linetype = "dashed", alpha = 0.35) +
  geom_hline(yintercept = -log10(0.05), color = "grey5", linetype = "dashed", alpha = 0.35) +
  theme_classic(base_size = 20) +
  theme(
    axis.title.y = element_text(face = "bold", margin = margin(0, 20, 0, 0), size = rel(1.1), color = 'black'),
    axis.title.x = element_text(face = "bold", margin = margin(20, 0, 0, 0), size = rel(1.1), color = 'black',),
    axis.text = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5, color = "black"),
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.title = element_blank() 
  ) +
  guides(color = guide_legend(override.aes = list(size = 5)))  # bigger legend dots

#Print the plot
print(cac3_volcano_plot)

#Save Plot
ggsave("./cac3_volcano_plot_FINAL.pdf", plot=cac3_volcano_plot, width = 12, height = 9, unit="in",  dpi=400)

##set-7
#Assign categories with NS prioritized
set7_vol$category <- case_when(
  set7_vol$log2FoldChange > -1 & set7_vol$log2FoldChange < 1 ~ "NS",
  set7_vol$NCU %in% Prc2targets$V11 ~ "PRC2",
  set7_vol$log2FoldChange > 1 & set7_vol$padj < 0.05 ~ "Upregulated",
  set7_vol$log2FoldChange < -1 & set7_vol$padj < 0.05 ~ "Downregulated",
  TRUE ~ "NS"
)

#Reorder so PRC2 is plotted last (on top)
set7_vol <- set7_vol[order(set7_vol$category == "PRC2"), ]

#Plot
set7_volcano_plot <- ggplot(set7_vol, aes(
  x = log2FoldChange,
  y = -log10(padj),
  color = category
)) +
  geom_point(size = 3, alpha = 0.85) +
  scale_color_manual(
    values = c(
      "Upregulated" = "red",
      "Downregulated" = "blue",
      "PRC2" = "orange2",
      "NS" = "grey"
    ),
    breaks = c("Upregulated", "Downregulated", "PRC2"), # only show these in legend
    labels = c("Upregulated", "Downregulated", "PRC2")
  ) +
  labs(
    x = expression(paste("log"[2], "FoldChange(set-7/WT)")),
    y = expression(paste("-log"[10], "(p-value)")),
    title = "set-7 vs WT"
  ) + 
  scale_x_continuous(breaks = breaks_width(3), minor_breaks = breaks_width(1)) + 
  scale_y_continuous(trans = pseudo_log_trans(base = 10, sigma = 2),
                     breaks = c(0,2,5,10,20,50,100,200),
                     limits = c(0, NA)) + 
  coord_cartesian(ylim = c(0, 200)) +
  geom_vline(xintercept = c(-1, 1), color = "grey5", linetype = "dashed", alpha = 0.35) +
  geom_hline(yintercept = -log10(0.05), color = "grey5", linetype = "dashed", alpha = 0.35) +
  theme_classic(base_size = 20) +
  theme(
    axis.title.y = element_text(face = "bold", margin = margin(0, 20, 0, 0), size = rel(1.1), color = 'black'),
    axis.title.x = element_text(face = "bold", margin = margin(20, 0, 0, 0), size = rel(1.1), color = 'black',),
    axis.text = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5, color = "black"),
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.title = element_blank() 
  ) +
  guides(color = guide_legend(override.aes = list(size = 5)))  # bigger legend dots

#Print the plot
print(set7_volcano_plot)

#Save Plot
ggsave("./set7_volcano_plot_FINAL.pdf", plot=set7_volcano_plot, width = 12, height = 9, unit="in",  dpi=400)

##cac-1_suz-12
#Assign categories with NS prioritized
cac1_suz12_vol$category <- case_when(
  cac1_suz12_vol$log2FoldChange > -1 & cac1_suz12_vol$log2FoldChange < 1 ~ "NS",
  cac1_suz12_vol$NCU %in% Prc2targets$V11 ~ "PRC2",
  cac1_suz12_vol$log2FoldChange > 1 & cac1_suz12_vol$padj < 0.05 ~ "Upregulated",
  cac1_suz12_vol$log2FoldChange < -1 & cac1_suz12_vol$padj < 0.05 ~ "Downregulated",
  TRUE ~ "NS"
)

#Reorder so PRC2 is plotted last (on top)
cac1_suz12_vol <- cac1_suz12_vol[order(cac1_suz12_vol$category == "PRC2"), ]

#Plot
library(ggplot2)
library(scales)  # breaks_width(), pseudo_log_trans()

cac1_suz12_volcano_plot <- ggplot(cac1_suz12_vol, aes(
  x = log2FoldChange,
  y = -log10(padj),
  color = category
)) +
  geom_point(size = 3, alpha = 0.85) +
  scale_color_manual(
    values = c(
      "Upregulated" = "red",
      "Downregulated" = "blue",
      "PRC2" = "orange2",
      "NS" = "grey"
    ),
    breaks = c("Upregulated", "Downregulated", "PRC2"), # only show these in legend
    labels = c("Upregulated", "Downregulated", "PRC2")
  ) +
  labs(
    x = expression(paste("log"[2], "FoldChange(cac-1_suz-12/WT)")),
    y = expression(paste("-log"[10], "(p-value)")),
    title = "cac-1_suz-12 vs WT"
  ) + 
  scale_x_continuous(breaks = breaks_width(3), minor_breaks = breaks_width(1)) + 
  scale_y_continuous(trans = pseudo_log_trans(base = 10, sigma = 2),
                     breaks = c(0,2,5,10,20,50,100,200,300),
                     limits = c(0, NA)) + 
  coord_cartesian(ylim = c(0, 300)) +
  geom_vline(xintercept = c(-1, 1), color = "grey5", linetype = "dashed", alpha = 0.35) +
  geom_hline(yintercept = -log10(0.05), color = "grey5", linetype = "dashed", alpha = 0.35) +
  theme_classic(base_size = 20) +
  theme(
    axis.title.y = element_text(face = "bold", margin = margin(0, 20, 0, 0), size = rel(1.1), color = 'black'),
    axis.title.x = element_text(face = "bold", margin = margin(20, 0, 0, 0), size = rel(1.1), color = 'black',),
    axis.text = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5, color = "black"),
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.title = element_blank() 
  ) +
  guides(color = guide_legend(override.aes = list(size = 5)))  # bigger legend dots

#Print the plot
print(cac1_suz12_volcano_plot)

ggsave("./cac1_suz12_volcano_plot_FINAL.pdf", plot=cac1_suz12_volcano_plot, width = 12, height = 9, unit="in",  dpi=400)

```

# Now, I want to make Venn-diagrams to compare the number of genes that are DE across all the CAF-1 mutants & set-7 for K27 genes
```{r Make Venn Diagrams, echo=FALSE}
#read in all Upregulated genes
cac1UP <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac1__alpha0.05__UP_genes.csv")
cac2UP <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac2__alpha0.05__UP_genes.csv")
cac3UP <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac3__alpha0.05__UP_genes.csv")
set7UP <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/set7__alpha0.05__UP_genes.csv")

cac1DOWN <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac1__alpha0.05__DOWN_genes.csv")
cac2DOWN <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac2__alpha0.05__DOWN_genes.csv")
cac3DOWN <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac3__alpha0.05__DOWN_genes.csv")
set7DOWN <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/set7__alpha0.05__DOWN_genes.csv")

cac1_K27 <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac1__alpha0.05__K27_UP_genes.csv")
cac2_K27 <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac2__alpha0.05__K27_UP_genes.csv")
cac3_K27 <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac3__alpha0.05__K27_UP_genes.csv")
set7_K27 <- read.csv("../CAF-1_RNA-seq_Analysis/csv_files/set7__alpha0.05__K27_UP_genes.csv")

#Venn Diagram w/ All Upregulated genes in CAF-1 mutants
venn.diagram(
  x = list(cac1UP$Attributes, cac2UP$Attributes, cac3UP$Attributes),
  category.names = c("cac1", "cac2 ", "cac3"),
  filename = 'CAF-1_UP_venn_diagramm.png',
  col=c("#8da0cb", "#66c2a5", "#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#66c2a5",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

#Make euler plot inputing values gathered from venn.diagram(). This will provide a proportional venn-diagram
CAF1_up_fit = euler(c("A" = 100, "B" = 218, "C" = 472, "A&B" = 254, "A&C" = 74, "B&C" =110, "A&B&C" = 695))
CAF1_up_eulerplot <- plot(CAF1_up_fit)

#Save file
ggsave(filename = "CAF-1_All_UP_euler.pdf", plot = CAF1_up_eulerplot, width = 4, height = 4, dpi = 600)

#Venn Diagram w/ All Downregulated genes in CAF-1 mutants
venn.diagram(
  x = list(cac1DOWN$Attributes, cac2DOWN$Attributes, cac3DOWN$Attributes),
  category.names = c("cac1", "cac2 ", "cac3"),
  filename = 'CAF-1_DOWN_venn_diagramm.png',
  col=c("#8da0cb", "#66c2a5", "#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#66c2a5",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

#Make euler plot inputing values gathered from venn.diagram(). This will provide a proportional venn-diagram
CAF1_down_fit = euler(c("A" = 91, "B" = 87, "C" = 448, "A&B" = 67, "A&C" = 57, "B&C" = 100, "A&B&C" = 200))
CAF1_down_eulerplot <- plot(CAF1_down_fit)

#Save file
ggsave(filename = "CAF-1_All_DOWN_euler.pdf", plot = CAF1_down_eulerplot, width = 4, height = 4, dpi = 600)

#Making Venn for All Upregulated K27 genes in CAF-1 mutants
venn.diagram(
  x = list(cac1_K27$Attributes, cac2_K27$Attributes, cac3_K27$Attributes),
  category.names = c("cac1", "cac2 ", "cac3"),
  filename = 'CAF-1_venn_diagramm_UP_K27_genes.png',
  col=c("#8da0cb", "#66c2a5", "#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#66c2a5",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

#plotting euler
CAF1_K27_up_fit = euler(c("A" = 7, "B" = 22, "C" = 77, "A&B" = 20, "A&C" = 20, "B&C" = 21, "A&B&C" = 202))
CAF1_K27_up_eulerplot <- plot(CAF1_K27_up_fit)

#Save file
ggsave(filename = "CAF-1_euler_K27_UP.pdf", plot = CAF1_K27_up_eulerplot, width = 4, height = 4, dpi = 600)

#Making Venn for All Upregulated genes in cac-1/2 and set-7 mutants
venn.diagram(
  x = list(cac1_K27$Attributes, cac2_K27$Attributes, set7_K27$Attributes),
  category.names = c("cac1", "cac2 ", "set7"),
  filename = 'cac1-2_set7_venn_diagramm_UP_K27_genes.png',
  col=c("#8da0cb", "#66c2a5", "#756bb1"),
  fill = c(alpha("#8da0cb",0.25), alpha("#66c2a5",0.25), alpha("#756bb1",0.25)),
  output=TRUE)

#plotting euler
cac1_2_set7_fit = euler(c("A" = 13, "B" = 12, "C" = 104, "A&B" = 56, "A&C" = 14, "B&C" = 31, "A&B&C" = 166))
cac1_2_set7_eulerplot <- plot(cac1_2_set7_fit)

#Save file
ggsave(filename = "cac1-2_set7_euler_All_UP.pdf", plot = cac1_2_set7_eulerplot, width = 4, height = 4, dpi = 600)
```

#Now, comparing genes that are upregulated in cac-1/2 and lose h3k27me3 and/or h3k36me3
```{r, venn test}
#combining RNA
library(dplyr)

# Stack then keep the first occurrence of each Attributes
cac1_suz12_K27UPcombine <- bind_rows(
  mutate(cac1_RNA_K27, .source = "cac1_RNA_K27"),
  mutate(suz12_RNA_K27, .source = "suz12_RNA_K27")
) %>%
  # if you want cac1UP to win on collisions, keep this order
  distinct(Attributes, .keep_all = TRUE) %>%
  select(-.source)

#only prc2 targets
cacupcombinek27 = subset(cacUPcombine, cacUPcombine$Attributes %in% Prc2targets$V10)

#Venn Diagram w/ All Upregulated genes in CAF-1 mutants
venn.diagram(
  x = list(cacupcombinek27$Attributes, K27lost$gene_id, K36lost$gene_id),
  category.names = c("Upregulated", "K27 lost ", "K36 lost"),
  filename = 'CAF-1_UP_K27K36loss_diagram.png',
  col=c("#8da0cb", "#66c2a5", "#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#66c2a5",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

#Make euler plot inputing values gathered from venn.diagram(). This will provide a proportional venn-diagram
CAF1_up_K27K36_fit = euler(c("A" = 126, "B" = 39, "C" = 10, "A&B" = 98, "A&C" = 19, "B&C" = 14, "A&B&C" = 49))
CAF1_up_K27K36_eulerplot <- plot(CAF1_up_K27K36_fit)

set7_RNA_K27 = read.csv("../CAF-1_RNA-seq_Analysis/csv_files/set7__alpha0.05__K27_UP_genes.csv")
suz12_RNA_K27 = read.csv("../CAF-1_RNA-seq_Analysis/csv_files/suz12__alpha0.05__K27_UP_genes.csv")
ash1_RNA_K27 = read.csv("../CAF-1_RNA-seq_Analysis/csv_files/ash1__alpha0.05__K27_UP_genes.csv")
cac1_RNA_K27 = read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac1__alpha0.05__K27_UP_genes.csv")
cac1_cac2RNA_K27 = read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac1_cac2__alpha0.05__K27_UP_genes.csv")
cac1_suz12RNA_K27 = read.csv("../CAF-1_RNA-seq_Analysis/csv_files/cac1_suz12__alpha0.05__K27_UP_genes.csv")

##################################
#FIGURE 2
#Patterns of genes that lose H3K27me3 in CAF-1 mutants
venn.diagram(
  x = list(K27lost_code$gene_id, cacupcombinek27$Attributes),
  category.names = c("K27_lost", "K27lost_CAF_UP"),
  filename = "K27lost_CAF-1_UP.png",
  col=c("#8da0cb","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)


CAF1_K27_lost_UP_fit = euler(c("A" = 43, "B" = 145, "A&B" = 147))
CAF1_K27_lost_UP_eulerplot <- plot(CAF1_K27_lost_UP_fit)
ggsave(filename = "CAF1_K27_lost_UP.pdf", plot = CAF1_K27_lost_UP_eulerplot, width = 4, height = 4, dpi = 600)

K27lost_cac1cac2UP = subset(K27lost_code, K27lost_code$gene_id %in% cacupcombinek27$Attributes)
K27lost_set7UP = subset(K27lost_code, K27lost_code$gene_id %in% set7_RNA_K27$Attributes)

venn.diagram(
  x = list(K27lost_set7UP$gene_id, K27lost_cac1cac2UP$gene_id),
  category.names = c("K27lost_set7_UP", "K27lost_CAF_UP"),
  filename = "K27lost_set7UP_CAF1UP.png",
  col=c("#8da0cb","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

CAF1_K27_lost_set7_UP_fit = euler(c("A" = 13, "B" = 48, "A&B" = 99))
CAF1_K27_lost_set7_UP_eulerplot <- plot(CAF1_K27_lost_set7_UP_fit)
ggsave(filename = "CAF1_K27_lost_set7_UP.pdf", plot = CAF1_K27_lost_set7_UP_eulerplot, width = 4, height = 4, dpi = 600)
##################################


###################################################
#FIGURE 3
#K27 lost ash1 K27/K36 lost caf1
venn.diagram(
  x = list(K27lostash1$gene_id, K27lost$gene_id),
  category.names = c("ash1 K27 lost", "CAF K27 lost"),
  filename = "ASH1_CAF1_K27K36lost.png",
  col=c("#8da0cb","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

CAF1_ASH1_K27lost_fit = euler(c("A" = 310, "B" = 48, "A&B" = 152))
CAF1_ASH1_K27lost_eulerplot <- plot(CAF1_ASH1_K27lost_fit)
ggsave(filename = "CAF1_ASH1_K27lost.pdf", plot = CAF1_ASH1_K27lost_eulerplot <- plot(CAF1_ASH1_K27lost_fit), width = 4, height = 4, dpi = 600)

#K27 genes upregulated ash1 caf1
venn.diagram(
  x = list(ash1_RNA_K27$Attributes, cacupcombinek27$Attributes),
  category.names = c("ash1 UP", "CAF UP"),
  filename = 'ASH1_CAF1_K27_UP_genes.png',
  col=c("#8da0cb","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

CAF1_ASH1_K27_UP_fit = euler(c("A" = 148, "B" = 36, "A&B" = 256))
CAF1_ASH1_K27_UP_eulerplot <- plot(CAF1_ASH1_K27_UP_fit)
ggsave(filename = "CAF-1_UP_ASH1_UP.pdf", plot = CAF1_ASH1_K27_UP_eulerplot <- plot(CAF1_ASH1_K27_UP_fit), width = 4, height = 4, dpi = 600)

#K27 genes upregulated ash1 set7
venn.diagram(
  x = list(ash1_RNA_K27$Attributes, set7_RNA_K27$Attributes),
  category.names = c("ash1 UP", "set7 UP"),
  filename = 'ASH1_set7_K27_UP_genes.png',
  col=c("#8da0cb","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

SET7_ASH1_K27_UP_fit = euler(c("A" = 172, "B" = 83, "A&B" = 232))
SET7_ASH1_K27_UP_eulerplot <- plot(SET7_ASH1_K27_UP_fit)
ggsave(filename = "SET7_UP_ASH1_UP.pdf", plot = SET7_ASH1_K27_UP_eulerplot <- plot(SET7_ASH1_K27_UP_fit), width = 4, height = 4, dpi = 600)

#K27 genes upregulated CAF1 K36 K27 lost
venn.diagram(
  x = list(K36lost_code$gene_id, K27lost_code$gene_id, cacupcombinek27$Attributes),
  category.names = c("K36lost", "K27 lost", "CAF UP"),
  filename = 'K36lost_K27lost_CAF_UP_genes.png',
  col=c("#8da0cb","#756bb1","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#756bb1",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

K36lost_K27lost_CAF_UP_fit = euler(c("A" = 37, "B" = 7, "C" = 127, "A&B" = 6, "A&C" = 102, "B&C" = 18, "A&B&C" = 45))
K36lost_K27lost_CAF_UP_eulerplot <- plot(K36lost_K27lost_CAF_UP_fit)
ggsave(filename = "K36lost_K27lost_CAF_UP.pdf", plot = K36lost_K27lost_CAF_UP_eulerplot, width = 4, height = 4, dpi = 600)

#K36 lost genes CAF-1 & set-7
venn.diagram(
  x = list(K36lost$gene_id, K36lost_set7$gene_id),
  category.names = c("CAF-1 K36 lost", "set-7 K36 lost"),
  filename = "CAF1_SET7_K36lost.png",
  col=c("#8da0cb","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

CAF1_SET7_K36lost_fit = euler(c("A" = 79, "B" = 28, "A&B" = 13))
CAF1_SET7_K36lost_eulerplot <- plot(CAF1_SET7_K36lost_fit)
ggsave(filename = "CAF1_SET7_K36lost.pdf", plot = CAF1_SET7_K36lost_eulerplot, width = 4, height = 4, dpi = 600)
####################################################

#####################################################
#FIGURE 4
###main
venn.diagram(
  x = list(Kac_K4_CAF_Gain_code$gene_id, cacupcombinek27$Attributes, K27lost_code$gene_id),
  category.names = c("CAF gained", "CAF UP", "K27lost"),
  filename = 'CAF1_Kac_K4_gained_Upreg_genes.png',
  col=c("#8da0cb","#756bb1","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#756bb1",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

Kac_K4_gain_K27lost_CAF_Up_fit = euler(c("A" = 13, "B" = 29, "C" = 131, "A&B" = 14, "A&C" = 14, "B&C" = 71, "A&B&C" = 76))
Kac_K4_gain_K27lost_CAF_Up_eulerplot <- plot(Kac_K4_gain_K27lost_CAF_Up_fit)
ggsave(filename = "Kac_K4_gain_K27lost_CAF_Up.pdf", plot = Kac_K4_gain_K27lost_CAF_Up_eulerplot, width = 4, height = 4, dpi = 600)

venn.diagram(
  x = list(Kac_Gain_CAF$gene_id, Kac_Gain_set7$gene_id),
  category.names = c("CAF gained", "SET7 gained"),
  filename = 'CAF1_set7_Kac_gained_genes.png',
  col=c("#8da0cb","#756bb1"),
  fill = c(alpha("#8da0cb",0.25), alpha("#756bb1",0.25)),
  output=TRUE)

CAF1_SET7_K27lost_Kac_Gain_fit = euler(c("A" = 92, "B" = 37, "A&B" = 12))
CAF1_SET7_K27lost_Kac_Gain_eulerplot <- plot(CAF1_SET7_K27lost_Kac_Gain_fit)
ggsave(filename = "CAF1_SET7_Kac_Gain.pdf", plot = CAF1_SET7_K27lost_Kac_Gain_eulerplot, width = 4, height = 4, dpi = 600)


venn.diagram(
  x = list(K4_Gain_CAF$gene_id, K4_Gain_set7$gene_id),
  category.names = c("CAF gained", "SET7 gained"),
  filename = 'CAF1_set7_K4me2_gained_genes.png',
  col=c("#8da0cb","#756bb1"),
  fill = c(alpha("#8da0cb",0.25), alpha("#756bb1",0.25)),
  output=TRUE)

CAF1_SET7_K27lost_K4_Gain_fit = euler(c("A" = 2, "B" = 26, "A&B" = 22))
CAF1_SET7_K27lost_K4_Gain_eulerplot <- plot(CAF1_SET7_K27lost_K4_Gain_fit)
ggsave(filename = "CAF1_SET7_K4_Gain.pdf", plot = CAF1_SET7_K27lost_K4_Gain_eulerplot, width = 4, height = 4, dpi = 600)

###supplement
venn.diagram(
  x = list(Kac_Gain_CAF$gene_id, K36lost$gene_id, K27lost$gene_id),
  category.names = c("CAF Kac gained", "CAF K36 lost", "CAF K27 lost"),
  filename = 'CAF1_Kac_gained_K27me3_K36me3_lost_genes.png',
  col=c("#8da0cb","#756bb1","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#756bb1",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

CAF1_Kac_Gain_K36_K27_lost_fit = euler(c("A" = 13, "B" = 21, "C" = 94, "A&B" = 5, "A&C" = 49, "B&C" = 20, "A&B&C" = 37))
CAF1_Kac_Gain_K36_K27_lost_eulerplot <- plot(CAF1_Kac_Gain_K36_K27_lost_fit)
ggsave(filename = "CAF1_Kac_Gain_K36_K27_lost.pdf", plot = CAF1_Kac_Gain_K36_K27_lost_eulerplot, width = 4, height = 4, dpi = 600)


#venn.diagram(
#  x = list(K4_Gain_CAF$gene_id, K27lost$gene_id),
#  category.names = c("CAF K4 gained", "CAF K27 lost"),
#  filename = 'CAF1_K4gain_K27lost_genes.png',
#  col=c("#8da0cb","#756bb1"),
#  fill = c(alpha("#8da0cb",0.25), alpha("#756bb1",0.25)),
#  output=TRUE)


venn.diagram(
  x = list(Kac_K4_cac1_Gain$gene_id, Kac_K4_cac2_Gain$gene_id),
  category.names = c("cac-1 gained", "cac-2 gained"),
  filename = 'cac1_cac2_Kac_K4_gained_genes.png',
  col=c("#8da0cb","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

cac1_cac2_Kac_K4_gained_fit = euler(c("A" = 61, "B" = 2, "A&B" = 60))
cac1_cac2_Kac_K4_gained_eulerplot <- plot(cac1_cac2_Kac_K4_gained_fit)
ggsave(filename = "cac1_cac2_Kac_K4_gained.pdf", plot = cac1_cac2_Kac_K4_gained_eulerplot, width = 4, height = 4, dpi = 600)

####################################################


#######################################################
#FIGURE 5
###Double Mutants
venn.diagram(
  x = list(cac1_cac2RNA_K27$Attributes, cacupcombinek27$Attributes),
  category.names = c("dub UP", "CAF UP"),
  filename = 'CAF1_cac1_2_K27_UP_genes.png',
  col=c("#8da0cb","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

CAF1_Double_K27_UP_fit = euler(c("A" = 182, "B" = 8, "A&B" = 284))
CAF1_Double_K27_UP_eulerplot <- plot(CAF1_Double_K27_UP_fit)
ggsave(filename = "CAF-1_Double_UP.pdf", plot = CAF1_Double_K27_UP_eulerplot, width = 4, height = 4, dpi = 600)

venn.diagram(
  x = list(cac1_suz12_K27UPcombine$Attributes, cac1_suz12RNA_K27$Attributes),
  category.names = c("singles up", "double up"),
  filename = 'cac1suz12_singles_double_mutant_K27_UP_genes.png',
  col=c("#8da0cb","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)

CAF1_suz12_Double_K27_UP_fit = euler(c("A" = 146, "B" = 5, "A&B" = 354))
CAF1_suz12_Double_K27_UP_eulerplot <- plot(CAF1_suz12_Double_K27_UP_fit)
ggsave(filename = "CAF-1_suz12_Double_UP.pdf", plot = CAF1_suz12_Double_K27_UP_eulerplot, width = 4, height = 4, dpi = 600)

#######################################################
#FIGURE 6
###Double Mutants
venn.diagram(
  x = list(K4_genes_gain_suz12$gene_id, K4_genes_gain_cac1$gene_id, K4_cac1_suz12_gain$gene_id),
  category.names = c("suz12 gain", "cac1 gain", "double gain"),
  filename = 'cac1suz12_K4_gain_genes.png',
  col=c("#8da0cb","#fc8d62","#66c2a5"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25), alpha("#66c2a5",0.25)),
  output=TRUE)

CAF1_suz12_Double_K4_fit = euler(c("A" = 0, "B" = 0, "C" = 7, "A&B" = 0, "A&C" = 5, "B&C" = 41, "A&B&C" = 9))
CAF1_suz12_Double_K4_eulerplot <- plot(CAF1_suz12_Double_K4_fit)
ggsave(filename = "CAF1_suz12_Double_K4.pdf", plot = CAF1_suz12_Double_K4_eulerplot, width = 4, height = 4, dpi = 600)

venn.diagram(
  x = list(Kac_Gain_suz12$gene_id, Kac_Gain_cac1$gene_id, Kac_cac1_suz12_gain$gene_id),
  category.names = c("suz12 gain", "cac1 gain", "double gain"),
  filename = 'cac1suz12_Kac_gain_genes.png',
  col=c("#8da0cb","#fc8d62","#66c2a5"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25), alpha("#66c2a5",0.25)),
  output=TRUE)

CAF1_suz12_Double_Kac_fit = euler(c("A" = 13, "B" = 33, "C" = 33, "A&B" = 1, "A&C" = 14, "B&C" = 14, "A&B&C" = 2))
CAF1_suz12_Double_Kac_eulerplot <- plot(CAF1_suz12_Double_Kac_fit)
ggsave(filename = "CAF1_suz12_Double_Kac.pdf", plot = CAF1_suz12_Double_Kac_eulerplot, width = 4, height = 4, dpi = 600)


###Tests
venn.diagram(
  x = list(ash1_RNA_K27$Attributes, set7_RNA_K27$Attributes, K27lost$gene_id),
  category.names = c("ash1_up", "set7_up", "K27 lost"),
  filename = 'ASH1UP_SET7UP_K27lostCAF.png',
  col=c("#8da0cb","#fc8d62","#66c2a5"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25), alpha("#66c2a5",0.25)),
  output=TRUE)

venn.diagram(
  x = list(ash1_RNA_K27$Attributes, set7_RNA_K27$Attributes, K36lost$gene_id),
  category.names = c("ash1_up", "set7_up", "K36 lost"),
  filename = 'ASH1UP_SET7UP_K36lostCAF.png',
  col=c("#8da0cb","#fc8d62","#756bb1"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25), alpha("#756bb1",0.25)),
  output=TRUE)

venn.diagram(
  x = list(ash1_RNA_K27$Attributes, cacupcombinek27$Attributes, CAF_K27_K36_Lost$gene_id),
  category.names = c("ash1_up", "cac_up", "K36 lost"),
  filename = 'ASH1UP_CAC1UP_K36K27lostCAF.png',
  col=c("#8da0cb","#fc8d62","#756bb1"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25), alpha("#756bb1",0.25)),
  output=TRUE)

#euler
CAF1_up_ASH1_up_K27K36_fit = euler(c("A" = 116, "B" = 17, "C" = 31, "A&B" = 109, "A&C" = 32, "B&C" = 19, "A&B&C" = 147))
CAF1_up_ASH1_up_K27K36_eulerplot <- plot(CAF1_up_ASH1_up_K27K36_fit)

#Save file
ggsave(filename = "CAF-1_Up_ASH1_Up_K27K36_loss.pdf", plot = CAF1_up_ASH1_up_K27K36_eulerplot <- plot(CAF1_up_ASH1_up_K27K36_fit), width = 4, height = 4, dpi = 600)

###CAF-1 ectopic genes
venn.diagram(
  x = list(cac1DOWN$Attributes, CAF1_ectopic_K27_genes_CODE$gene_id),
  category.names = c("down", "ectopic"),
  filename = 'CAF-1_ectopic_gene_downreg.png',
  col=c("#8da0cb","#fc8d62"),
  fill = c(alpha("#8da0cb",0.25), alpha("#fc8d62",0.25)),
  output=TRUE)


```

#Lastly, I want to make an IGV track for visualizing DE data with ChIP data
```{r Make IGV Track, echo=FALSE}
#calling list of files
##Note: this will only work if the ONLY csv's in this directory are the ones containing the Data you want on the track (i.e. you need to move any other csv's from this directory, or move the ones you want on the track to a new one and change the path)
list_of_files <- list.files(path = "../Research",
                            recursive = FALSE,
                            pattern = ".csv$")

induced_genes <- readr::read_csv(list_of_files, id = "file_name")

##convert to wide format
l2fc_data <- data.frame(pivot_wider(data = induced_genes, id_cols = "NCU", names_from = "file_name", values_from = "log2FoldChange"))

#assigning row names and deleting extra column
rownames(l2fc_data) <- l2fc_data[,1]
l2fc_data <- l2fc_data[,-1]

#assigning NA values to 0
l2fc_data[is.na(l2fc_data)] <- 0

#tibble NCU names back to first column
l2fc <- tibble::rownames_to_column(l2fc_data, "VALUE")

#read in file w/ all genes
ngenes = read.csv("../csv_files/neurospora_genes_edit.csv")

#Make IGV Track
##Note: only genes with DE in one or more strains will display data in the track
combine <- transform(merge(ngenes, l2fc, by = "VALUE"))
track = combine %>% select("SequenceID", "FeatureStart", "FeatureEnd", "VALUE", "asf1__alpha0.05__DE_absLFC_ge_1p0.csv", "atrx__alpha0.05__DE_absLFC_ge_1p0.csv", "cac1__alpha0.05__DE_absLFC_ge_1p0.csv", "cac2__alpha0.05__DE_absLFC_ge_1p0.csv", "cac3__alpha0.05__DE_absLFC_ge_1p0.csv", "naf1__alpha0.05__DE_absLFC_ge_1p0.csv", "naf2__alpha0.05__DE_absLFC_ge_1p0.csv", "set7__alpha0.05__DE_absLFC_ge_1p0.csv", "suz12__alpha0.05__DE_absLFC_ge_1p0.csv", "ash1__alpha0.05__DE_absLFC_ge_1p0.csv", "cac1_cac2__alpha0.05__DE_absLFC_ge_1p0.csv", "cac1_suz12__alpha0.05__DE_absLFC_ge_1p0.csv")

#write track to file
write.table(track, file="RNAseq_CAF-1_Paper_Final.igv", sep="\t", row.names = FALSE, quote = FALSE)
```

#END